---
title: "Big Cubes"
author: "Lasse Hjorth Madsen"
date: today 
format: 
  html:
    fig-width: 4
    fig-height: 4
    df-print: paged
toc: true
toc-expand: true
toc-depth: 4
editor: source
execute:
  echo: false
  warning: false
  cache: true
---

```{r setup}
#| cache: false
library(tidyverse)

devtools::load_all(path = "../../bglab")

met <- get_met()

theme_set(
  theme_minimal() +
  theme(
    axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
    panel.grid.major.y = element_blank()
    )
  )
```

```{r helpers}
remove_stuff <- function(text, leading_breaks = 4) {
  text |>
    read_lines() |>
    str_remove("^ {1,6}") |>
    str_remove("Confidence.+") |>
    str_remove("Duration.+") |>
    str_remove("Rollout¹    ") |> 
    str_replace_all("        ", "  ") |> 
    paste(collapse = "\n") |>  
    str_replace_all("\n\n", "\n") |> 
    cat(rep("\n", leading_breaks), ... = _, sep = "")
}
```

```{r takepoints}
tps <- expand_grid(x = 3:25, cube = c(1, 2, 4, 8, 16)) |> 
  rowwise() |>  
  mutate(tp = tp(x, x, cube, met),
         single_win_value = mwc(x - cube, x , met) - mwc(x ,x , met)) |>  
  ungroup() |>  
  filter(tp > 0, cube < x) |> 
  arrange(-tp)
```

## Why this?

This is to investigate take points when the cube level gets high and the match score is tied.

## Example

Here's one example from a recent match:

```{r}
#| column: page-right
#| layout-ncol: 2

ggboard("XGID=-CC--A------------------h-:2:-1:-1:00:3:3:0:17:10") +
  labs(title = "Black on roll, cube action?")

"Analyzed in Rollout
Redouble/Take
  Player Winning Chances:   76,10% (G:0,00% B:0,00%)
  Opponent Winning Chances: 23,90% (G:0,00% B:0,00%)

Cubeful Equities:
       No redouble:     +0,809 (-0,168)
       Redouble/Take:   +0,977
       Redouble/Pass:   +1,000 (+0,023)
" |> remove_stuff() |> cat()
```

In the match, I was playing White and passed this cube, which was a very small mistake. I expected the take point to be a bit higher than for money, but apparently that is not the case.

My thinking was, that points get progressively more valuable the closer you get to the end of a match: At double match point you have 50% chance of winning, an actual win get you to 100%, so a point is worth 50 percentage points. At 17-away, 17-away a single point win get you from 50% to 54% so is worth just 4 percentage points. At 5-away, 5-away a point is worth about 8 percentage points.

So in a sense doubling more than doubles the stakes. A bit like 'doubling' from 4 to not 8 but perhaps 9 points. That should make you *more* reluctant to take, I thought. 

An actual calculation shows that the take point when doubled from 4 to 8 at 13-away, 13-away is
`r round(tp(14, 14, 4, met) * 100, 1)`% -- practically the same as for money, if anything a bit *lower*.

That agrees with the numbers in the position above; White can take with just 23.9%, since he gains a bit from being able to double Black out efficiently when they get to a two-roll position. 

So exactly where did my logic go wrong?

The thing is, as the cube get bigger, the value of the points *currently* at stake also increases, so doubling does not necessarily "more that doubles" the stakes as I thought. Here's a table to illustrate:


```{r table}
tps |> 
  select(-tp) |> 
  mutate(single_win_value = single_win_value * 100) |> 
  filter(x %in% 9:21, cube %in% c(4, 8)) |> 
  arrange(x) |>
  rename(`Match length` = x) |> 
  mutate(cube = paste("Value of a", cube, "point win")) |> 
  pivot_wider(names_from = cube, values_from = single_win_value) |>  
  mutate(Ratio = c_across(3) / c_across(2)) |> 
  mutate(across(starts_with("Value"), \(x) round(x, 1)),
         across(Ratio, \(x) round(x, 2))) |> 
  kableExtra::kable()
```

We see that the value of both a 4-point win and an 8-point win increases as the match gets shorter, but the *ratio* between the two remains approximately the same, just about 2. So doubling from 4 to 8 does in fact double the stakes, not more than that.

Let's plot the take points for more scores and more cube values:

```{r plot}
#| column: page-right
#| fig-width: 6

tps |> 
  ggplot(aes(x = x, y = tp, color = factor(cube))) + 
  geom_point() + 
  geom_line() +
  geom_hline(yintercept = 0.25, color = "darkgray") +
  scale_x_continuous(breaks = 3:25) +
  scale_y_continuous(breaks = seq(0, 0.4, 0.02)) +
  labs(title = "Cubeless, gammonless take points at even score",
       subtitle = "by match length and cube value",
       color = "Cube value",
       x = "Match length",
       y = "Take point") +
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(), 
        panel.grid.major.y = element_line())
```

Near the end of the match the take points can be a bit funny, as we know. When both players need 4 points, it's hard to take a 4-cube (take point `r round(tp(4, 4, 2, met) * 100, 0)`%). Here, doubling from 2 to 4 does indeed "more that doubles" the stakes. 

At the other extreme, at 5-away, 5-away, it's easy to take an 8-cube, because the alternative, dropping, would leave you with just `r round(met[5,1] * 100, 1)`% winning chances.

At longer matches, when the score is tied, cubes at the 1-, 2- or 4-levels behaves just like for money, as we might expect.

8- and 16-cubes are different, however. These take-point are quite a bit lower that 25% (unless the match is *very* long). There intuition here is that conceding 8 or 16 points would leave you in a pretty desperate situaition anyway, so it's not something one should do easily. 

