---
title: "Hit or point decisions"
author: "Lasse Hjorth Madsen"
date: today 
format: 
  html:
    fig-width: 4
    fig-height: 4
toc: true
toc-expand: true
toc-depth: 2
editor: source
execute:
  echo: false
  warning: false
  cache: true
---

```{r setup}
#| cache: false
 
library(tidyverse)
devtools::load_all(path = "../../bglab")

remove_stuff <- function(text) {
  text |>
    read_lines() |>
    str_remove("^ {1,6}") |>
    str_remove("Confidence.+") |>
    str_remove("Duration.+") |>
    str_remove("Rollout¹    ") |> 
    str_replace_all("        ", "  ") |> 
    paste(collapse = "\n") |>  
    str_replace_all("\n\n", "\n") 
}
```

## Why this?

This is a collection of checker play problems from the middle game, where the decision is between hitting or making a point. This can often be a difficult choice, so I thought it would be useful to take a closer look. 

We have five main positions, each with several variants, all are rolled out 1296 times; moves and cube decisions at 3-ply.

## Position 1

In the first two contrasting variants, hitting with 21 is clearly right, for all the usual reasons: It gains in the race, gains initiative, cuts down on Black's good numbers, leaves bad 6s from the bar. 

#### Variation A

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a----E-D-a-eD---c-e-A--A-:0:0:1:21:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    13/10*                       eq:+0,138
      Player:   52,70% (G:15,93% B:0,75%)
      Opponent: 47,30% (G:11,61% B:0,60%)
      Confidence: ±0,014 (+0,124..+0,152) - [100,0%]
      Duration: 7 minutes 11 seconds

    2. Rollout¹    24/21                        eq:+0,018 (-0,120)
      Player:   49,69% (G:11,01% B:0,34%)
      Opponent: 50,31% (G:7,46% B:0,27%)
      Confidence: ±0,013 (+0,006..+0,031) - [0,0%]
      Duration: 6 minutes 40 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

In the second position, either of the pointing plays are a lot better that hitting. 

#### Variation B

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a----E-Da--eD---c-eA---A-:0:0:1:31:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    8/5 6/5                      eq:+0,239
      Player:   54,79% (G:17,12% B:0,74%)
      Opponent: 45,21% (G:9,91% B:0,37%)
      Confidence: ±0,014 (+0,225..+0,253) - [100,0%]
      Duration: 6 minutes 11 seconds

    2. Rollout¹    24/20                        eq:+0,186 (-0,053)
      Player:   53,47% (G:11,71% B:0,29%)
      Opponent: 46,53% (G:5,87% B:0,24%)
      Confidence: ±0,012 (+0,173..+0,198) - [0,0%]
      Duration: 6 minutes 40 seconds

    3. Rollout²    13/9*                        eq:+0,134 (-0,105)
      Player:   52,80% (G:16,03% B:0,75%)
      Opponent: 47,20% (G:12,05% B:0,59%)
      Confidence: ±0,014 (+0,120..+0,148) - [0,0%]
      Duration: 21 minutes 25 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

Why the difference?

Let's try to ignore the even stronger pointing play of 8/5 6/5 in variation B, and just compare the advanced anchor plays to the hitting plays.

Clearly, the 20-point is stronger than the 21-point, so that's a long term reason why pointing is more attractive in variation B than in variation A: The point we get to make is better.

There're also a related tactical reason: Leaving a blot on the 20-point in variation B gives Black nice, loose return hits with aces: 21, 31, 41, and 51 from the bar. In variation A the same aces are not nearly as comfortable after hitting. (The corresponding loose hits with 2s are not as clear and happy for Black.) 

Another way of putting it: Hitting in variation A gives Black *both* poor 6s and aces from the bar; in variation B it is only 6s that are uncomfortable. 

For these reasons, the defensive anchor works better that hitting i variation A but not in variation B.

As mentioned, B offers White an even stronger option: 8/5 6 /5. That's better because the 5-point is also a great point to make, *and* it's harder since it's not slotted like the 20-point is. So it makes sense to take the opportunity now, while the need for defense is not super urgent. 

## Position 2
#### Variation A

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-b---BD-Ca--cD---bbcA-bA--:0:0:1:13:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    23/20 6/5                    eq:+0,557
      Player:   62,59% (G:18,02% B:0,71%)
      Opponent: 37,41% (G:6,52% B:0,28%)
      Confidence: ±0,014 (+0,544..+0,571) - [100,0%]
      Duration: 5 minutes 01 second

    2. Rollout¹    13/9*                        eq:+0,414 (-0,143)
      Player:   59,06% (G:20,49% B:1,62%)
      Opponent: 40,94% (G:11,74% B:0,61%)
      Confidence: ±0,015 (+0,400..+0,429) - [0,0%]
      Duration: 5 minutes 11 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

In position 2A, making the point with 23/20 6/5 is far ahead of the hitting play, 13/9\*, for several reasons. Of course, the anchor is super strong, and a permanent asset. But that's always the case, and often hitting is stronger anyway. In this position specifically: 

- The non-hitting play is nice and smooth
- The non-hitting play leaves no blots, avoiding 54 and 63 fly shots 
- Hitting lets Black make an advanced anchor with 43 and 32
- Hitting lets Black return hit loose with 41 and 31 (much like in the previous position)

#### Variation B

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-b---BD-Ca--cD---bbcA-bA--:0:0:1:24:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    8/4 6/4                      eq:+0,472
      Player:   60,45% (G:21,72% B:1,62%)
      Opponent: 39,55% (G:12,06% B:0,47%)
      Confidence: ±0,015 (+0,456..+0,487) - [100,0%]
      Duration: 4 minutes 28 seconds

    2. Rollout¹    23/21 13/9*                  eq:+0,423 (-0,049)
      Player:   59,54% (G:20,22% B:1,44%)
      Opponent: 40,46% (G:11,35% B:0,58%)
      Confidence: ±0,015 (+0,408..+0,437) - [0,0%]
      Duration: 5 minutes 38 seconds

    3. Rollout¹    13/11 13/9*                  eq:+0,399 (-0,072)
      Player:   58,74% (G:21,36% B:1,74%)
      Opponent: 41,26% (G:12,68% B:0,93%)
      Confidence: ±0,017 (+0,382..+0,417) - [0,0%]
      Duration: 5 minutes 33 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

Same position, different roll, now with a choice of making the 4-point or hitting. Pointing is still right, but with a smaller margin, since the 4-point is not as urgent, and not as important as the 20-point anchor in the previous position.

#### Variation C

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-b---BD-C-a-cD---bbcA-bA--:0:0:1:23:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    23/21 13/10*                 eq:+0,457
      Player:   60,45% (G:19,87% B:1,57%)
      Opponent: 39,55% (G:11,44% B:0,58%)
      Confidence: ±0,019 (+0,439..+0,476) - [99,1%]
      Duration: 5 minutes 33 seconds

    2. Rollout¹    13/11 13/10*                 eq:+0,424 (-0,033)
      Player:   59,11% (G:21,10% B:1,79%)
      Opponent: 40,89% (G:12,23% B:0,86%)
      Confidence: ±0,020 (+0,404..+0,444) - [0,8%]
      Duration: 5 minutes 09 seconds

    3. Rollout¹    23/20 8/6                    eq:+0,420 (-0,037)
      Player:   59,11% (G:17,36% B:0,61%)
      Opponent: 40,89% (G:7,19% B:0,30%)
      Confidence: ±0,013 (+0,407..+0,433) - [0,1%]
      Duration: 5 minutes 07 seconds

    4. Rollout¹    13/10* 10/8                  eq:+0,403 (-0,054)
      Player:   59,11% (G:19,61% B:1,33%)
      Opponent: 40,89% (G:11,19% B:0,47%)
      Confidence: ±0,015 (+0,388..+0,418) - [0,0%]
      Duration: 5 minutes 07 seconds

    5. Rollout¹    23/20 13/11                  eq:+0,397 (-0,060)
      Player:   58,87% (G:16,95% B:0,67%)
      Opponent: 41,13% (G:7,76% B:0,28%)
      Confidence: ±0,012 (+0,385..+0,410) - [0,0%]
      Duration: 5 minutes 19 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

Almost the same position; this time the anchoring play is a little awkward, and that is enough to make hitting correct, by a small amount. 

#### Variation D

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-b---BD-BA-acD---bbcA-bA--:0:0:1:25:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    9/4 6/4                      eq:+0,423
      Player:   59,38% (G:20,66% B:1,37%)
      Opponent: 40,62% (G:12,40% B:0,56%)
      Confidence: ±0,016 (+0,408..+0,439) - [73,8%]
      Duration: 3 minutes 17 seconds

    2. Rollout¹    20/15 13/11*                 eq:+0,416 (-0,008)
      Player:   60,01% (G:20,37% B:1,57%)
      Opponent: 39,99% (G:13,02% B:0,89%)
      Confidence: ±0,017 (+0,399..+0,433) - [26,1%]
      Duration: 4 minutes 03 seconds

    3. Rollout¹    13/8 13/11*                  eq:+0,389 (-0,035)
      Player:   58,76% (G:20,90% B:1,73%)
      Opponent: 41,24% (G:12,24% B:0,85%)
      Confidence: ±0,018 (+0,371..+0,407) - [0,1%]
      Duration: 4 minutes 07 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

Variation D is a photo finish: The pointing play cleans up a blot and is super smooth, the hitting play unstacks the mid and is very flexible. The hit has all the usual benefits, while the inner point takes a lot of the sting out of any loose hit from Black. 

Bottom line: The two plays are basically equal.

#### Variation E

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-b---BDAB-a-cD---bbcA-bA--:0:0:1:23:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    23/20 7/5                    eq:+0,515
      Player:   61,14% (G:19,33% B:0,67%)
      Opponent: 38,86% (G:6,94% B:0,29%)
      Confidence: ±0,015 (+0,500..+0,530) - [100,0%]
      Duration: 3 minutes 51 seconds

    2. Rollout¹    7/4 6/4                      eq:+0,435 (-0,080)
      Player:   59,61% (G:21,34% B:1,48%)
      Opponent: 40,39% (G:12,48% B:0,56%)
      Confidence: ±0,015 (+0,420..+0,449) - [0,0%]
      Duration: 3 minutes 18 seconds

    3. Rollout¹    13/10* 7/5                   eq:+0,402 (-0,112)
      Player:   58,53% (G:21,25% B:1,69%)
      Opponent: 41,47% (G:11,84% B:0,68%)
      Confidence: ±0,016 (+0,386..+0,418) - [0,0%]
      Duration: 3 minutes 53 seconds

    4. Rollout¹    13/10* 10/8                  eq:+0,361 (-0,154)
      Player:   57,81% (G:19,84% B:1,53%)
      Opponent: 42,19% (G:11,58% B:0,55%)
      Confidence: ±0,016 (+0,345..+0,377) - [0,0%]
      Duration: 4 minutes 28 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

In variation E, if White plays 7/5 with the two, the best 3 is 23/20, making the achor, for the same reasons as seen earlier. The 20-point is simple very valueable, and Black would like to hit loose there, so making it takes priority. The play is also quite flexible, so beats hitting. 

Between the two pointing plays, the anchor wins for the same reason as before: It is the better and more urgent point to make; the 4-point is likely to come later.

#### Variation F

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-b---BD-C--acD---bbc-AbA--:0:0:1:23:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    23/20 13/11*                 eq:+0,464
      Player:   60,81% (G:19,52% B:1,50%)
      Opponent: 39,19% (G:11,05% B:0,49%)
      Confidence: ±0,016 (+0,447..+0,480) - [99,8%]
      Duration: 4 minutes 08 seconds

    2. Rollout¹    23/21 13/10                  eq:+0,433 (-0,030)
      Player:   59,47% (G:18,57% B:0,73%)
      Opponent: 40,53% (G:8,19% B:0,34%)
      Confidence: ±0,013 (+0,421..+0,446) - [0,2%]
      Duration: 3 minutes 44 seconds

    3. Rollout¹    13/10 13/11*                 eq:+0,404 (-0,059)
      Player:   59,19% (G:20,57% B:1,50%)
      Opponent: 40,81% (G:12,62% B:0,86%)
      Confidence: ±0,015 (+0,389..+0,420) - [0,0%]
      Duration: 3 minutes 52 seconds

    4. Rollout¹    13/11* 11/8                  eq:+0,398 (-0,066)
      Player:   58,75% (G:19,49% B:1,25%)
      Opponent: 41,25% (G:11,55% B:0,50%)
      Confidence: ±0,016 (+0,381..+0,414) - [0,0%]
      Duration: 3 minutes 46 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

When the point to make is a slightly inferior anchor, the scales tip again and hitting is best, by a small amount, noting that 23/21 13/11\* makes a bid for the better anchor *while* hitting. 

## Position 3
#### Variation A

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a-Ba-DaB--AcB---bbe-A--BA:0:0:1:54:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    Bar/20 24/20                 eq:-0,171
      Player:   45,64% (G:12,63% B:0,53%)
      Opponent: 54,36% (G:14,37% B:0,59%)
      Confidence: ±0,016 (-0,187..-0,155) - [100,0%]
      Duration: 5 minutes 39 seconds

    2. Rollout¹    Bar/20 11/7*                 eq:-0,209 (-0,038)
      Player:   44,86% (G:13,48% B:0,63%)
      Opponent: 55,14% (G:16,50% B:1,14%)
      Confidence: ±0,016 (-0,225..-0,193) - [0,0%]
      Duration: 5 minutes 40 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation B

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a-Ba-DaB--AcB---bbe-A--BA:0:0:1:34:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    Bar/21 11/8                  eq:-0,272
      Player:   43,03% (G:11,62% B:0,43%)
      Opponent: 56,97% (G:14,27% B:0,72%)
      Confidence: ±0,014 (-0,285..-0,258) - [80,9%]
      Duration: 5 minutes 11 seconds

    2. Rollout¹    Bar/22 11/7*                 eq:-0,284 (-0,013)
      Player:   43,33% (G:12,70% B:0,61%)
      Opponent: 56,67% (G:17,81% B:1,21%)
      Confidence: ±0,018 (-0,302..-0,266) - [12,5%]
      Duration: 5 minutes 48 seconds

    3. Rollout¹    Bar/21 24/21                 eq:-0,286 (-0,015)
      Player:   42,86% (G:11,61% B:0,49%)
      Opponent: 57,14% (G:15,66% B:0,69%)
      Confidence: ±0,016 (-0,302..-0,271) - [6,6%]
      Duration: 5 minutes 34 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation C

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a-Ba-DaBA--cB---bbe-A--BA:0:0:1:45:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    Bar/21 9/4*                  eq:-0,181
      Player:   45,16% (G:14,43% B:0,70%)
      Opponent: 54,84% (G:15,73% B:0,92%)
      Confidence: ±0,017 (-0,198..-0,164) - [97,8%]
      Duration: 5 minutes 27 seconds

    2. Rollout¹    Bar/20 24/20                 eq:-0,205 (-0,024)
      Player:   44,93% (G:12,75% B:0,55%)
      Opponent: 55,07% (G:14,61% B:0,66%)
      Confidence: ±0,016 (-0,221..-0,189) - [2,2%]
      Duration: 5 minutes 45 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation D

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a-Ba-D-B-aAcB---bbe-A--BA:0:0:1:15:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    Bar/20 11/10*                eq:-0,198
      Player:   45,33% (G:13,20% B:0,74%)
      Opponent: 54,67% (G:16,62% B:1,20%)
      Confidence: ±0,016 (-0,214..-0,182) - [100,0%]
      Duration: 5 minutes 46 seconds

    2. Rollout¹    Bar/20 21/20                 eq:-0,258 (-0,060)
      Player:   43,49% (G:12,44% B:0,51%)
      Opponent: 56,51% (G:15,10% B:0,89%)
      Confidence: ±0,018 (-0,276..-0,240) - [0,0%]
      Duration: 5 minutes 45 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

## Position 4
#### Variation A

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=----b-E-C---aD--be-eB---A-:0:0:1:16:0:0:0:0:10
"
ggboard(id)

"    1. Rollout¹    24/18 13/12*                 eq:+0,054
      Player:   51,13% (G:9,31% B:0,40%)
      Opponent: 48,87% (G:9,61% B:0,31%)
      Confidence: ±0,014 (+0,040..+0,068) - [100,0%]
      Duration: 5 minutes 25 seconds

    2. Rollout¹    13/7 13/12*                  eq:-0,010 (-0,064)
      Player:   50,54% (G:10,34% B:0,44%)
      Opponent: 49,46% (G:12,37% B:0,60%)
      Confidence: ±0,016 (-0,026..+0,006) - [0,0%]
      Duration: 5 minutes 33 seconds

    3. Rollout¹    13/7 8/7                     eq:-0,019 (-0,073)
      Player:   49,12% (G:9,23% B:0,37%)
      Opponent: 50,88% (G:10,49% B:0,31%)
      Confidence: ±0,014 (-0,032..-0,005) - [0,0%]
      Duration: 4 minutes 32 seconds

    4. Rollout¹    13/12* 12/6                  eq:-0,041 (-0,095)
      Player:   48,74% (G:9,61% B:0,31%)
      Opponent: 51,26% (G:9,98% B:0,37%)
      Confidence: ±0,013 (-0,055..-0,028) - [0,0%]
      Duration: 4 minutes 55 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation B

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=----b-E-C--AaC--be-eB---A-:0:0:1:16:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    11/5 6/5                     eq:+0,093
      Player:   51,96% (G:10,47% B:0,35%)
      Opponent: 48,04% (G:9,26% B:0,29%)
      Confidence: ±0,013 (+0,080..+0,106) - [99,9%]
      Duration: 4 minutes 17 seconds

    2. Rollout¹    24/18 13/12*                 eq:+0,062 (-0,030)
      Player:   51,75% (G:9,27% B:0,35%)
      Opponent: 48,25% (G:9,71% B:0,37%)
      Confidence: ±0,013 (+0,049..+0,076) - [0,1%]
      Duration: 5 minutes 13 seconds

    3. Rollout¹    13/12* 12/6                  eq:-0,026 (-0,119)
      Player:   49,23% (G:9,90% B:0,34%)
      Opponent: 50,77% (G:10,23% B:0,35%)
      Confidence: ±0,013 (-0,039..-0,013) - [0,0%]
      Duration: 4 minutes 48 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation C

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=---b--E-C--a-D--be-eB---A-:0:0:1:24:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    24/20 13/11*                 eq:+0,218
      Player:   55,42% (G:10,96% B:0,47%)
      Opponent: 44,58% (G:9,21% B:0,32%)
      Confidence: ±0,015 (+0,203..+0,232) - [92,7%]
      Duration: 5 minutes 13 seconds

    2. Rollout¹    8/4 6/4                      eq:+0,203 (-0,015)
      Player:   54,90% (G:12,57% B:0,50%)
      Opponent: 45,10% (G:10,05% B:0,38%)
      Confidence: ±0,013 (+0,190..+0,216) - [7,3%]
      Duration: 4 minutes 30 seconds

    3. Rollout¹    13/11* 11/7                  eq:+0,105 (-0,112)
      Player:   52,73% (G:12,78% B:0,55%)
      Opponent: 47,27% (G:12,03% B:0,50%)
      Confidence: ±0,015 (+0,091..+0,120) - [0,0%]
      Duration: 5 minutes 36 seconds

    4. Rollout¹    13/9 13/11*                  eq:+0,091 (-0,127)
      Player:   52,50% (G:13,20% B:0,59%)
      Opponent: 47,50% (G:13,01% B:0,67%)
      Confidence: ±0,016 (+0,075..+0,107) - [0,0%]
      Duration: 5 minutes 22 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation D

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=----b-E-C---aE--be-eB-----:0:0:1:13:0:0:0:0:1"
ggboard(id)

"    1. Rollout¹    13/10 13/12*                 eq:+0,107
      Player:   52,96% (G:9,55% B:0,33%)
      Opponent: 47,04% (G:9,00% B:0,31%)
      Confidence: ±0,013 (+0,095..+0,120) - [97,4%]
      Duration: 9 minutes 10 seconds

    2. Rollout¹    13/12* 12/9                  eq:+0,087 (-0,021)
      Player:   52,20% (G:9,28% B:0,33%)
      Opponent: 47,80% (G:8,99% B:0,34%)
      Confidence: ±0,014 (+0,072..+0,101) - [1,7%]
      Duration: 9 minutes 05 seconds

    3. Rollout¹    13/12* 6/3                   eq:+0,084 (-0,023)
      Player:   52,12% (G:9,34% B:0,39%)
      Opponent: 47,88% (G:9,77% B:0,35%)
      Confidence: ±0,015 (+0,070..+0,099) - [0,9%]
      Duration: 9 minutes 48 seconds

    4. Rollout¹    8/5 6/5                      eq:+0,080 (-0,028)
      Player:   51,25% (G:8,53% B:0,30%)
      Opponent: 48,75% (G:7,17% B:0,21%)
      Confidence: ±0,012 (+0,068..+0,092) - [0,1%]
      Duration: 7 minutes 42 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation E

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=----b-F-B---aD--be-eB---A-:0:0:1:13:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    13/12* 6/3                   eq:-0,072
      Player:   48,37% (G:10,66% B:0,42%)
      Opponent: 51,63% (G:13,49% B:0,54%)
      Confidence: ±0,015 (-0,087..-0,056) - [78,7%]
      Duration: 13 minutes 30 seconds

    2. Rollout¹    8/5 6/5                      eq:-0,080 (-0,008)
      Player:   47,75% (G:10,86% B:0,41%)
      Opponent: 52,25% (G:12,35% B:0,53%)
      Confidence: ±0,013 (-0,093..-0,066) - [21,1%]
      Duration: 10 minutes 57 seconds

    3. Rollout¹    24/21 13/12*                 eq:-0,101 (-0,030)
      Player:   47,74% (G:9,00% B:0,28%)
      Opponent: 52,26% (G:10,85% B:0,38%)
      Confidence: ±0,017 (-0,118..-0,084) - [0,2%]
      Duration: 12 minutes 37 seconds

    4. Rollout¹    13/12* 12/9                  eq:-0,105 (-0,033)
      Player:   47,46% (G:10,25% B:0,37%)
      Opponent: 52,54% (G:12,58% B:0,46%)
      Confidence: ±0,015 (-0,120..-0,090) - [0,0%]
      Duration: 12 minutes 31 seconds

    5. Rollout¹    13/10 13/12*                 eq:-0,111 (-0,040)
      Player:   47,47% (G:10,25% B:0,40%)
      Opponent: 52,53% (G:13,36% B:0,55%)
      Confidence: ±0,015 (-0,126..-0,096) - [0,0%]
      Duration: 12 minutes 38 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation F

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=----b-D-C---aE--be-eB---A-:0:0:1:13:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    8/5 6/5                      eq:+0,098
      Player:   52,47% (G:11,29% B:0,38%)
      Opponent: 47,53% (G:11,82% B:0,37%)
      Confidence: ±0,014 (+0,084..+0,112) - [96,1%]
      Duration: 12 minutes 26 seconds

    2. Rollout¹    13/10 13/12*                 eq:+0,075 (-0,023)
      Player:   51,85% (G:11,18% B:0,47%)
      Opponent: 48,15% (G:12,97% B:0,49%)
      Confidence: ±0,021 (+0,055..+0,096) - [3,9%]
      Duration: 14 minutes 27 seconds

    3. Rollout¹    24/21 13/12*                 eq:+0,025 (-0,073)
      Player:   50,83% (G:9,78% B:0,37%)
      Opponent: 49,17% (G:11,41% B:0,37%)
      Confidence: ±0,016 (+0,010..+0,041) - [0,0%]
      Duration: 14 minutes 11 seconds

    4. Rollout¹    13/12* 12/9                  eq:+0,007 (-0,091)
      Player:   50,73% (G:11,02% B:0,44%)
      Opponent: 49,27% (G:13,15% B:0,55%)
      Confidence: ±0,017 (-0,010..+0,024) - [0,0%]
      Duration: 14 minutes 09 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation G

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=----b-D-C---aF--be-eB-----:0:0:1:13:0:0:0:0:1"
ggboard(id)

"    1. Rollout¹    13/10 13/12*                 eq:+0,139
      Player:   53,55% (G:9,67% B:0,38%)
      Opponent: 46,45% (G:9,97% B:0,35%)
      Confidence: ±0,015 (+0,124..+0,155) - [100,0%]
      Duration: 10 minutes 57 seconds

    2. Rollout¹    13/12* 12/9                  eq:+0,083 (-0,056)
      Player:   52,21% (G:9,68% B:0,38%)
      Opponent: 47,79% (G:10,14% B:0,33%)
      Confidence: ±0,013 (+0,070..+0,096) - [0,0%]
      Duration: 10 minutes 45 seconds

    3. Rollout¹    8/5 6/5                      eq:+0,054 (-0,085)
      Player:   50,80% (G:8,65% B:0,33%)
      Opponent: 49,20% (G:8,80% B:0,25%)
      Confidence: ±0,014 (+0,041..+0,068) - [0,0%]
      Duration: 9 minutes 08 seconds

    4. Rollout¹    13/12* 6/3                   eq:+0,035 (-0,104)
      Player:   51,30% (G:9,36% B:0,33%)
      Opponent: 48,71% (G:11,12% B:0,37%)
      Confidence: ±0,016 (+0,019..+0,052) - [0,0%]
      Duration: 11 minutes 42 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

## Position 5
#### Variation A

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a----EaC--AdD---d-e---AA-:0:0:1:56:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    13/7* 6/1*                   eq:+0,098
      Player:   51,88% (G:17,84% B:0,49%)
      Opponent: 48,12% (G:13,89% B:0,76%)
      Confidence: ±0,014 (+0,083..+0,112) - [100,0%]
      Duration: 4 minutes 51 seconds

    2. Rollout¹    24/18 23/18                  eq:+0,002 (-0,096)
      Player:   49,85% (G:9,63% B:0,30%)
      Opponent: 50,15% (G:9,14% B:0,32%)
      Confidence: ±0,012 (-0,010..+0,014) - [0,0%]
      Duration: 5 minutes 23 seconds

    3. Rollout¹    13/8 13/7*                   eq:-0,020 (-0,118)
      Player:   49,28% (G:14,77% B:0,63%)
      Opponent: 50,72% (G:14,47% B:0,71%)
      Confidence: ±0,015 (-0,036..-0,005) - [0,0%]
      Duration: 5 minutes 07 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation B

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=-a----EaD---dD---d-e---AA-:0:0:1:56:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    24/18 23/18                  eq:+0,117
      Player:   53,05% (G:9,04% B:0,26%)
      Opponent: 46,95% (G:7,06% B:0,25%)
      Confidence: ±0,011 (+0,106..+0,128) - [85,3%]
      Duration: 4 minutes 49 seconds

    2. Rollout¹    13/7* 6/1*                   eq:+0,108 (-0,009)
      Player:   51,95% (G:17,39% B:0,49%)
      Opponent: 48,05% (G:12,98% B:0,62%)
      Confidence: ±0,014 (+0,094..+0,121) - [14,7%]
      Duration: 4 minutes 27 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation C

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=--a---EaD---dD---d-e---AA-:0:0:1:56:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    24/18 23/18                  eq:+0,148
      Player:   53,61% (G:9,59% B:0,29%)
      Opponent: 46,39% (G:7,46% B:0,24%)
      Confidence: ±0,012 (+0,135..+0,160) - [100,0%]
      Duration: 4 minutes 47 seconds

    2. Rollout¹    13/7* 7/2*                   eq:+0,065 (-0,083)
      Player:   50,99% (G:15,87% B:0,58%)
      Opponent: 49,01% (G:12,65% B:0,54%)
      Confidence: ±0,013 (+0,052..+0,079) - [0,0%]
      Duration: 4 minutes 22 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^

#### Variation D

```{r}
#| layout-ncol: 2
#| column: page-right

id <- "XGID=--a---EaD---dD---d-e--A-A-:0:0:1:46:0:0:0:0:10"
ggboard(id)

"    1. Rollout¹    8/2* 6/2                     eq:+0,197
      Player:   53,51% (G:18,40% B:0,57%)
      Opponent: 46,49% (G:10,13% B:0,45%)
      Confidence: ±0,013 (+0,183..+0,210) - [100,0%]
      Duration: 3 minutes 43 seconds

    2. Rollout¹    13/7* 6/2*                   eq:+0,161 (-0,035)
      Player:   53,09% (G:17,54% B:0,62%)
      Opponent: 46,91% (G:12,22% B:0,54%)
      Confidence: ±0,014 (+0,147..+0,176) - [0,0%]
      Duration: 4 minutes 36 seconds

    3. Rollout¹    24/18 22/18                  eq:+0,144 (-0,053)
      Player:   53,39% (G:9,69% B:0,26%)
      Opponent: 46,61% (G:7,33% B:0,21%)
      Confidence: ±0,012 (+0,132..+0,155) - [0,0%]
      Duration: 4 minutes 49 seconds
" |>  
  remove_stuff() |>  
  cat(sep = "")
```

^`r id`^
