---
title: "Exploring take points"
author: "Lasse Hjorth Madsen"
date: today 
format: html
toc: true
toc-expand: true
editor: visual
execute:
  echo: false
  warning: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(kableExtra)

devtools::load_all(path = "../../bglab")

theme_set(
  theme_minimal() +
    theme(
      legend.position = c(1, 1),
      legend.justification = c(1, 1),
      legend.background = element_rect(fill = "white", color = NA)
    )
)
```

## Why this?

What is the lowest take point you can possibly have, with reasonable gammon frequencies? Not surprisingly, this happens when you're trailing 17-away, 17-away, the opponent holds a 16-cube and double to 32. If you pass, you have about 1 percentage chance of winning, so if you got that many winning chances, you can take.

But what about all the other take points; which ones are the most extreme?

## A big plot

These are some assumed outcome probabilities, similar to the opening position:

```{r takepoints}
met <- get_met()
probs <- outcome_probs(c(50.0, 13.0, 0.5, 50.0, 13.0, 0.5))

tps <- expand_grid(x = 2:17, y = 2:17, cube = c(1, 2, 4, 8, 16)) %>% 
  rowwise() %>% 
  mutate(tp = tp_info(x, y, probs, cube, met)["tp_real"]) %>% 
  ungroup() %>% 
  filter(tp > 0, cube < x) %>% 
  mutate(score_label = paste0("-", x, ",-", y), tp) %>% 
  arrange(-tp)
```

```{r assumptions}
probs_table(probs) %>% 
  round(1) %>% 
  kable("html") %>%
  column_spec(2:5, width = "3.1cm")
```

Using those gammon and backgammon frequencies, here's a take point plot, for all possible scores and cube-values in a 17-point match. It's fun to explore and learn. Can you find that one purple 16-cube mentioned in the beginning?

```{r plot}
#| fig-height: 28

all_scores <- tps %>% 
  filter(cube == 1) %>% 
  arrange(-tp) 

score_order <- all_scores %>% 
  pull(score_label)

score_color <- all_scores %>%
  mutate(score_color = case_when(x > y ~ "darkred",
                                 x < y ~ "darkblue",
                                 TRUE ~ "black")) %>% 
  pull(score_color)

names(score_color) <- score_order

tps %>% 
  ggplot(aes(y = score_label, x = tp, group = score_label)) +
  geom_point(aes(color = factor(cube))) +
  geom_line(color = "grey") +
  scale_y_discrete(limits = score_order) +
  scale_x_continuous(breaks = seq(0, 1, 0.05), minor_breaks = NULL) +
  labs(title = "Take points in 17-point matches, with medium gammon frequencies",
       subtitle = "by score and cube level. First score is takers' score. Score colors indicate a lead/trail/tie",
       x = "Take point", y = "Score", color = "Cube level") +
  theme(axis.text.y = element_text(colour = score_color))
```
