---
title: "Middle game cube benchmark"
author: "Lasse Hjorth Madsen"
date: today 
format: html
toc: false
toc-expand: true
toc-depth: 4
editor: source
execute:
  echo: false
  warning: false
  cache: false
---

```{r setup}
library(tidyverse)
library(stringr)
library(kableExtra)
devtools::load_all(path = "../../bglab")
max_move <- 12
error_threshold <- 0.07
met <- get_met()

remove_stuff <- function(text) {
  text %>% 
  str_remove_all(" Winning Chances") %>% 
  str_remove("\nCubeless Equities.*\n") 
}
```

```{r subset_and_clean}
df <- bgmoves %>% 
  distinct(xgid, .keep_all = TRUE) %>%
  filter(move_no <= max_move, 
         score1 == 0, score2 == 0, 
         proper_ca == "Double, take") %>% 
  # We don't want both the double and the subsequent take decision:
  mutate(duplet = file == lag(file) & move_no - 1 == lag(move_no)) %>% 
  filter(is.na(duplet) | !duplet) %>% 
  rowwise() %>% 
  # Extract the size of pass and no double errors
  mutate(pass_error = read_lines(cube_eq)[6],
         no_double_error = read_lines(cube_eq)[7]) %>% 
  ungroup() %>% 
  mutate(across(c(pass_error, no_double_error), 
                ~ str_extract(., "\\(.+\\)$") %>% 
                  str_remove_all("\\(|\\)") %>% 
                  as.numeric())) %>% 
  filter(pass_error > error_threshold, no_double_error < -error_threshold) %>% 
  select(file, move_no, player1, player2, cube_eq, xgid)  
```

## Why this?

At [Backgammon Galaxy](play.backgammongalaxy.com), I play mostly 3-point matches, because this is the match length where getting a game is easiest. But correct cube action at the beginning of 3-point matches is different than from money, so I'm a bit concerned that playing so many 3-pointers will bias my cube strategy in longer matches.

To get a better feel for the difference, this is a collection of backgammon positions that GNU Backgammon 4-ply considered solid double/takes at 3-away, 3-away, but not necessarily for money and longer matches. The positions are all from the early middle game (here defined as move number `r max_move` at the most, counting both sides). By 'solid double/take' I mean, that both not doubling and passing would be at least a `r error_threshold` error.

The positions are extracted semi-automatically from my database of games, analyzed by GNU Backgammon. For that reason sometimes they are shown as checker plays -- this happens when one player failed to double. For the same reason, sometimes Black is on roll, sometimes White.

Finally, while GNU Backgammon at 4-ply thought all positions where clear double/take at 3-away, 3-away, the XG rollouts sometimes come to a different conclusion: One side has a pass, or a no-double.

## A bit of theory

Let's revisit how and why the cube actions are different at 3-away, 3-away than for money. The 3-point match equity table looks like this:

```{r}
met[1:3, 1:3] %>% round(3)
info <- tp_info(3, 3, c(0.5, 0, 0, 0.5, 0, 0 ), 1, met)
```

In the simple case, not considering gammons and recubes, when we pass an initial double we get to `r met[3,2] %>% round(3)` match winning chances, losing about 10 percentage points compared to when the game began. If we take, a loss is expensive, we get to `r met[3,1] %>% round(3)` match winning chances, an additional loss of about 15 percentage points.

So the *risk* from taking and winning is `r info["risk"] %>% round(3)`, and the potential *gain* from taking and winning is `r info["gain"] %>% round(3)` for a take point of `r info["tp_dead"] %>% round(3)`, based on the standard formula: $TP= \frac{Risk}{Gain + Risk}$. So a pure 4-roll position, for example, would be a pass.

After taking, the *gammon value* with the cube at 2, is `r gammon_value(3, 3, 2, met) %>% round(3)`, not much different from the money gammon value of 0.5, so adding gammons into the mix has about the same effect as in money game.

After taking, the opponent's take point for a future redouble, is just `r met[3,1] %>% round(3)` which is *higher* than for money, since the cube would be dead. That means that the cube is a little *more* valuable to us than for money, because we don't have to improve quite as much to use it efficiently.

For illustration purposes, let's assume gammon- and backgammon probabilities that are similar to the opening position for money, like this:

```{r}
probs <- outcome_probs(c(50, 14, 0.6, 50, 14, 0.6)) %>% `/`(100) 

probs %>% 
  probs_table() %>% 
  kable("html") %>%
  column_spec(2:5, width = "3.1cm") 
```

Then, putting it all together, we can calculate take points under different assumptions, and compare to money in a nice table like this:

```{r}
tp_tab <- tp_table(3, 3, probs, 1, met)

tp_tab %>% 
  mutate(across(where(is.numeric), \(x) round(x, digits = 3)),
         Difference = `3-away, 3-away take point` - `Money game take point`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

Sure enough, the table shows that the 'simple' take point (no gammons, no cube) that we just discussed, is significant higher at the score than for money. A little bit of the difference disappear when we introduce gammons; somewhat more when we introduce the cube.

The end result is that the take point at 3-away 3-away is about 1.5 percentage point higher than for money. It may not sound as much, but the difference is quite real, as the examples below will show.

## Double/take positions

Below are 21 positions rolled out at 3-away, 3-away and for money. Although many of the examples are double/take both at the score and for money, the side on roll is always more inclined to double at the match score than for money. The side not on roll, is always closer to a pass at the score than for money.

Studying the positions can be useful to reinforce one's impression of what it takes to have a sound double, and what a takeable position looks like -- for money and at 3-away, 3-away. 

### Position 1

```{r}
ggboard("XGID=---BBBC-B---bB--b-bcBc-c--:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2

"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   73,84% (G:3,02% B:0,10%)
  Opponent Winning Chances: 26,16% (G:2,68% B:0,08%)
Double/Take
  Player Winning Chances:   74,11% (G:3,04% B:0,07%)
  Opponent Winning Chances: 25,89% (G:2,71% B:0,08%)

Cubeless Equities: No Double=+0,482, Double=+1,224

Cubeful Equities:
       No double:     +0,871 (-0,129)
       Double/Take:   +1,053 (+0,053)
       Double/Pass:   +1,000

Best Cube action: Double / Pass

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   73,86% (G:3,02% B:0,04%)
  Opponent Winning Chances: 26,14% (G:2,52% B:0,08%)
Double/Take
  Player Winning Chances:   74,13% (G:2,95% B:0,04%)
  Opponent Winning Chances: 25,87% (G:2,52% B:0,09%)

Cubeless Equities: No Double=+0,482, Double=+0,973

Cubeful Equities:
       No double:     +0,754 (-0,095)
       Double/Take:   +0,849
       Double/Pass:   +1,000 (+0,151)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 2

```{r}
ggboard("XGID=-aa---EBB-AAbC---d-fa---A-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   68,34% (G:20,05% B:2,05%)
  Opponent Winning Chances: 31,66% (G:8,13% B:0,65%)
Double/Take
  Player Winning Chances:   68,54% (G:20,65% B:2,80%)
  Opponent Winning Chances: 31,46% (G:8,91% B:1,01%)

Cubeless Equities: No Double=+0,583, Double=+1,229

Cubeful Equities:
       No double:     +0,834 (-0,152)
       Double/Take:   +0,986
       Double/Pass:   +1,000 (+0,014)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   73,86% (G:3,02% B:0,04%)
  Opponent Winning Chances: 26,14% (G:2,52% B:0,08%)
Double/Take
  Player Winning Chances:   74,13% (G:2,95% B:0,04%)
  Opponent Winning Chances: 25,87% (G:2,52% B:0,09%)

Cubeless Equities: No Double=+0,482, Double=+0,973

Cubeful Equities:
       No double:     +0,754 (-0,095)
       Double/Take:   +0,849
       Double/Pass:   +1,000 (+0,151)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 3

```{r}
ggboard("XGID=-a-B--D-B---dD---dAdb--AA-:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   65,91% (G:24,59% B:2,02%)
  Opponent Winning Chances: 34,09% (G:7,59% B:0,48%)
Double/Take
  Player Winning Chances:   65,57% (G:25,68% B:4,39%)
  Opponent Winning Chances: 34,43% (G:8,14% B:0,72%)

Cubeless Equities: No Double=+0,615, Double=+1,225

Cubeful Equities:
       No double:     +0,799 (-0,097)
       Double/Take:   +0,896
       Double/Pass:   +1,000 (+0,104)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   65,97% (G:24,76% B:1,34%)
  Opponent Winning Chances: 34,03% (G:7,63% B:0,27%)
Double/Take
  Player Winning Chances:   66,16% (G:24,55% B:1,29%)
  Opponent Winning Chances: 33,84% (G:7,49% B:0,27%)

Cubeless Equities: No Double=+0,501, Double=+1,008

Cubeful Equities:
       No double:     +0,718 (-0,027)
       Double/Take:   +0,746
       Double/Pass:   +1,000 (+0,254)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 4

```{r}
ggboard("XGID=---BBBC-A---gC---a-cbb--AA:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   69,13% (G:13,40% B:0,47%)
  Opponent Winning Chances: 30,87% (G:6,28% B:0,25%)
Double/Take
  Player Winning Chances:   69,29% (G:13,98% B:0,63%)
  Opponent Winning Chances: 30,71% (G:6,63% B:0,40%)

Cubeless Equities: No Double=+0,496, Double=+1,157

Cubeful Equities:
       No double:     +0,793 (-0,113)
       Double/Take:   +0,906
       Double/Pass:   +1,000 (+0,094)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   69,19% (G:12,98% B:0,36%)
  Opponent Winning Chances: 30,81% (G:6,15% B:0,15%)
Double/Take
  Player Winning Chances:   69,37% (G:13,52% B:0,33%)
  Opponent Winning Chances: 30,63% (G:6,22% B:0,16%)

Cubeless Equities: No Double=+0,454, Double=+0,924

Cubeful Equities:
       No double:     +0,730
       Double/Take:   +0,721 (-0,010)
       Double/Pass:   +1,000 (+0,270)

Best Cube action: No double / Take
Percentage of wrong pass needed to make the double decision right: 3,4%

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 5

```{r}
ggboard("XGID=-a---CDBBa--dB---c-d-Ab-A-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   64,86% (G:25,92% B:1,39%)
  Opponent Winning Chances: 35,14% (G:8,94% B:0,78%)
Double/Take
  Player Winning Chances:   64,47% (G:27,45% B:2,06%)
  Opponent Winning Chances: 35,53% (G:9,55% B:1,00%)

Cubeless Equities: No Double=+0,571, Double=+1,178

Cubeful Equities:
       No double:     +0,762 (-0,095)
       Double/Take:   +0,857
       Double/Pass:   +1,000 (+0,143)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   64,69% (G:26,19% B:0,95%)
  Opponent Winning Chances: 35,31% (G:8,21% B:0,39%)
Double/Take
  Player Winning Chances:   64,76% (G:26,94% B:0,92%)
  Opponent Winning Chances: 35,24% (G:8,47% B:0,42%)

Cubeless Equities: No Double=+0,479, Double=+0,970

Cubeful Equities:
       No double:     +0,671 (-0,040)
       Double/Take:   +0,711
       Double/Pass:   +1,000 (+0,289)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 6

```{r}
ggboard("XGID=-a-BBBD-Aa--bB---c-cc--bB-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   62,31% (G:33,81% B:1,43%)
  Opponent Winning Chances: 37,69% (G:12,75% B:1,34%)
Double/Take
  Player Winning Chances:   62,46% (G:34,40% B:2,25%)
  Opponent Winning Chances: 37,54% (G:13,77% B:1,78%)

Cubeless Equities: No Double=+0,569, Double=+1,145

Cubeful Equities:
       No double:     +0,615 (-0,248)
       Double/Take:   +0,864
       Double/Pass:   +1,000 (+0,136)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   62,26% (G:33,32% B:0,93%)
  Opponent Winning Chances: 37,74% (G:12,35% B:0,95%)
Double/Take
  Player Winning Chances:   62,46% (G:34,17% B:0,99%)
  Opponent Winning Chances: 37,54% (G:12,63% B:1,02%)

Cubeless Equities: No Double=+0,455, Double=+0,929

Cubeful Equities:
       No double:     +0,542 (-0,091)
       Double/Take:   +0,634
       Double/Pass:   +1,000 (+0,366)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction.
" %>% 
  remove_stuff() %>% 
  cat()
```

### Position 7

```{r}
ggboard("XGID=aa----EBDA--bA---c-dbb-B--:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   67,45% (G:21,24% B:2,26%)
  Opponent Winning Chances: 32,55% (G:8,05% B:0,75%)
Double/Take
  Player Winning Chances:   66,74% (G:22,62% B:5,39%)
  Opponent Winning Chances: 33,26% (G:8,67% B:0,89%)

Cubeless Equities: No Double=+0,587, Double=+1,194

Cubeful Equities:
       No double:     +0,784 (-0,094)
       Double/Take:   +0,879
       Double/Pass:   +1,000 (+0,121)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   67,20% (G:21,36% B:1,24%)
  Opponent Winning Chances: 32,80% (G:7,89% B:0,39%)
Double/Take
  Player Winning Chances:   67,05% (G:21,78% B:1,40%)
  Opponent Winning Chances: 32,95% (G:7,85% B:0,36%)

Cubeless Equities: No Double=+0,487, Double=+0,982

Cubeful Equities:
       No double:     +0,703 (-0,026)
       Double/Take:   +0,729
       Double/Pass:   +1,000 (+0,271)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 8

```{r}
ggboard("XGID=---BaCC---C-cCab-c-c--b-A-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   63,32% (G:28,35% B:0,77%)
  Opponent Winning Chances: 36,68% (G:7,68% B:0,49%)
Double/Take
  Player Winning Chances:   63,76% (G:29,33% B:1,20%)
  Opponent Winning Chances: 36,24% (G:8,17% B:0,79%)

Cubeless Equities: No Double=+0,588, Double=+1,224

Cubeful Equities:
       No double:     +0,704 (-0,154)
       Double/Take:   +0,858
       Double/Pass:   +1,000 (+0,142)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   63,34% (G:28,13% B:0,58%)
  Opponent Winning Chances: 36,66% (G:6,89% B:0,35%)
Double/Take
  Player Winning Chances:   63,68% (G:28,77% B:0,67%)
  Opponent Winning Chances: 36,32% (G:7,35% B:0,37%)

Cubeless Equities: No Double=+0,481, Double=+0,982

Cubeful Equities:
       No double:     +0,632 (-0,053)
       Double/Take:   +0,686
       Double/Pass:   +1,000 (+0,314)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 9

```{r}
ggboard("XGID=-a--aBEBB---dBa--c-e----B-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   66,26% (G:28,97% B:2,59%)
  Opponent Winning Chances: 33,74% (G:8,04% B:0,66%)
Double/Take
  Player Winning Chances:   66,09% (G:30,01% B:4,18%)
  Opponent Winning Chances: 33,91% (G:8,92% B:1,15%)

Cubeless Equities: No Double=+0,692, Double=+1,340

Cubeful Equities:
       No double:     +0,831 (-0,169)
       Double/Take:   +1,031 (+0,031)
       Double/Pass:   +1,000

Best Cube action: Double / Pass

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   66,56% (G:27,73% B:1,42%)
  Opponent Winning Chances: 33,44% (G:7,32% B:0,36%)
Double/Take
  Player Winning Chances:   66,55% (G:28,81% B:1,45%)
  Opponent Winning Chances: 33,45% (G:7,60% B:0,39%)

Cubeless Equities: No Double=+0,546, Double=+1,107

Cubeful Equities:
       No double:     +0,740 (-0,104)
       Double/Take:   +0,844
       Double/Pass:   +1,000 (+0,156)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 10

```{r}
ggboard("XGID=-a-BB-D-AA--eC---c-cB-c---:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   66,26% (G:28,97% B:2,59%)
  Opponent Winning Chances: 33,74% (G:8,04% B:0,66%)
Double/Take
  Player Winning Chances:   66,09% (G:30,01% B:4,18%)
  Opponent Winning Chances: 33,91% (G:8,92% B:1,15%)

Cubeless Equities: No Double=+0,692, Double=+1,340

Cubeful Equities:
       No double:     +0,831 (-0,169)
       Double/Take:   +1,031 (+0,031)
       Double/Pass:   +1,000

Best Cube action: Double / Pass

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   66,20% (G:23,68% B:0,31%)
  Opponent Winning Chances: 33,80% (G:4,38% B:0,18%)
Double/Take
  Player Winning Chances:   66,51% (G:23,88% B:0,35%)
  Opponent Winning Chances: 33,49% (G:4,56% B:0,18%)

Cubeless Equities: No Double=+0,518, Double=+1,050

Cubeful Equities:
       No double:     +0,750 (-0,034)
       Double/Take:   +0,784
       Double/Pass:   +1,000 (+0,216)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 11

```{r}
ggboard("XGID=aa--ABBBA--AcD---cbe--A-A-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   66,56% (G:22,17% B:1,86%)
  Opponent Winning Chances: 33,44% (G:8,37% B:0,63%)
Double/Take
  Player Winning Chances:   66,63% (G:22,94% B:3,48%)
  Opponent Winning Chances: 33,37% (G:9,39% B:0,89%)

Cubeless Equities: No Double=+0,572, Double=+1,178

Cubeful Equities:
       No double:     +0,778 (-0,117)
       Double/Take:   +0,896
       Double/Pass:   +1,000 (+0,104)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   66,75% (G:22,02% B:0,99%)
  Opponent Winning Chances: 33,25% (G:7,98% B:0,40%)
Double/Take
  Player Winning Chances:   66,87% (G:22,39% B:1,02%)
  Opponent Winning Chances: 33,13% (G:8,18% B:0,40%)

Cubeless Equities: No Double=+0,481, Double=+0,971

Cubeful Equities:
       No double:     +0,721
       Double/Take:   +0,712 (-0,008)
       Double/Pass:   +1,000 (+0,279)

Best Cube action: No double / Take
Percentage of wrong pass needed to make the double decision right: 2,8%

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 12

```{r}
ggboard("XGID=-ba-AAE-Ca--dD---b-eA-----:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   65,19% (G:27,99% B:5,46%)
  Opponent Winning Chances: 34,81% (G:6,60% B:0,33%)
Double/Take
  Player Winning Chances:   64,35% (G:28,90% B:10,63%)
  Opponent Winning Chances: 35,65% (G:7,02% B:0,51%)

Cubeless Equities: No Double=+0,758, Double=+1,272

Cubeful Equities:
       No double:     +0,792 (-0,075)
       Double/Take:   +0,868
       Double/Pass:   +1,000 (+0,132)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction.
" %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   65,39% (G:27,20% B:2,73%)
  Opponent Winning Chances: 34,61% (G:6,62% B:0,25%)
Double/Take
  Player Winning Chances:   65,26% (G:27,59% B:2,93%)
  Opponent Winning Chances: 34,74% (G:6,89% B:0,32%)

Cubeless Equities: No Double=+0,538, Double=+1,077

Cubeful Equities:
       No double:     +0,743 (-0,041)
       Double/Take:   +0,784
       Double/Pass:   +1,000 (+0,216)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 13

```{r}
ggboard("XGID=-aa-aBDAB---dD---b-dBb----:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   66,66% (G:21,87% B:1,78%)
  Opponent Winning Chances: 33,34% (G:6,59% B:0,27%)
Double/Take
  Player Winning Chances:   66,45% (G:22,98% B:3,21%)
  Opponent Winning Chances: 33,55% (G:7,23% B:0,60%)

Cubeless Equities: No Double=+0,604, Double=+1,224

Cubeful Equities:
       No double:     +0,797 (-0,107)
       Double/Take:   +0,904
       Double/Pass:   +1,000 (+0,096)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   66,58% (G:21,88% B:1,17%)
  Opponent Winning Chances: 33,42% (G:6,14% B:0,22%)
Double/Take
  Player Winning Chances:   66,89% (G:21,88% B:1,19%)
  Opponent Winning Chances: 33,11% (G:6,25% B:0,22%)

Cubeless Equities: No Double=+0,498, Double=+1,007

Cubeful Equities:
       No double:     +0,734 (-0,034)
       Double/Take:   +0,769
       Double/Pass:   +1,000 (+0,231)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 14

```{r}
ggboard("XGID=---b--EDC---a---bbceA--AA-:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   67,40% (G:19,51% B:1,69%)
  Opponent Winning Chances: 32,60% (G:7,86% B:0,76%)
Double/Take
  Player Winning Chances:   67,48% (G:20,06% B:3,62%)
  Opponent Winning Chances: 32,52% (G:9,25% B:0,94%)

Cubeless Equities: No Double=+0,549, Double=+1,152

Cubeful Equities:
       No double:     +0,752 (-0,141)
       Double/Take:   +0,894
       Double/Pass:   +1,000 (+0,106)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   67,56% (G:19,29% B:1,00%)
  Opponent Winning Chances: 32,44% (G:7,87% B:0,35%)
Double/Take
  Player Winning Chances:   67,55% (G:19,48% B:1,05%)
  Opponent Winning Chances: 32,45% (G:7,89% B:0,35%)

Cubeless Equities: No Double=+0,472, Double=+0,948

Cubeful Equities:
       No double:     +0,671 (-0,034)
       Double/Take:   +0,705
       Double/Pass:   +1,000 (+0,295)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 15

```{r}
ggboard("XGID=-a-aBBDBBA--c----b-cc-b-B-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   64,07% (G:35,22% B:2,06%)
  Opponent Winning Chances: 35,93% (G:11,02% B:1,72%)
Double/Take
  Player Winning Chances:   64,39% (G:35,84% B:3,37%)
  Opponent Winning Chances: 35,61% (G:12,54% B:2,23%)

Cubeless Equities: No Double=+0,658, Double=+1,309

Cubeful Equities:
       No double:     +0,732 (-0,268)
       Double/Take:   +1,035 (+0,035)
       Double/Pass:   +1,000

Best Cube action: Double / Pass

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   64,28% (G:34,84% B:1,18%)
  Opponent Winning Chances: 35,72% (G:10,98% B:0,81%)
Double/Take
  Player Winning Chances:   64,54% (G:35,15% B:1,12%)
  Opponent Winning Chances: 35,46% (G:10,94% B:0,80%)

Cubeless Equities: No Double=+0,528, Double=+1,072

Cubeful Equities:
       No double:     +0,655 (-0,146)
       Double/Take:   +0,802
       Double/Pass:   +1,000 (+0,198)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 16

```{r}
ggboard("XGID=-a-aBAD-B--AdCa--b-ca-bAA-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   63,42% (G:30,48% B:4,32%)
  Opponent Winning Chances: 36,58% (G:10,19% B:1,62%)
Double/Take
  Player Winning Chances:   63,19% (G:30,89% B:7,60%)
  Opponent Winning Chances: 36,81% (G:11,54% B:1,57%)

Cubeless Equities: No Double=+0,645, Double=+1,150

Cubeful Equities:
       No double:     +0,750 (-0,094)
       Double/Take:   +0,844
       Double/Pass:   +1,000 (+0,156)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   63,29% (G:29,95% B:2,95%)
  Opponent Winning Chances: 36,71% (G:9,95% B:0,59%)
Double/Take
  Player Winning Chances:   63,34% (G:30,18% B:3,12%)
  Opponent Winning Chances: 36,66% (G:10,11% B:0,64%)

Cubeless Equities: No Double=+0,489, Double=+0,985

Cubeful Equities:
       No double:     +0,642 (-0,054)
       Double/Take:   +0,697
       Double/Pass:   +1,000 (+0,303)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 17

```{r}
ggboard("XGID=aA--A-CbC---dE-a-Abcb---A-:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   66,15% (G:24,46% B:2,49%)
  Opponent Winning Chances: 33,85% (G:7,38% B:0,49%)
Double/Take
  Player Winning Chances:   66,17% (G:25,61% B:5,05%)
  Opponent Winning Chances: 33,83% (G:8,13% B:0,86%)

Cubeless Equities: No Double=+0,633, Double=+1,253

Cubeful Equities:
       No double:     +0,820 (-0,110)
       Double/Take:   +0,930
       Double/Pass:   +1,000 (+0,070)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   66,33% (G:23,87% B:1,69%)
  Opponent Winning Chances: 33,67% (G:7,25% B:0,23%)
Double/Take
  Player Winning Chances:   66,56% (G:24,47% B:1,81%)
  Opponent Winning Chances: 33,44% (G:7,38% B:0,24%)

Cubeless Equities: No Double=+0,507, Double=+1,036

Cubeful Equities:
       No double:     +0,753 (-0,031)
       Double/Take:   +0,784
       Double/Pass:   +1,000 (+0,216)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 18

```{r}
ggboard("XGID=-baB--D-BAA-bC---abc-B-bb-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   66,97% (G:24,43% B:2,58%)
  Opponent Winning Chances: 33,03% (G:10,30% B:0,56%)
Double/Take
  Player Winning Chances:   67,02% (G:24,87% B:4,92%)
  Opponent Winning Chances: 32,98% (G:11,48% B:0,61%)

Cubeless Equities: No Double=+0,605, Double=+1,194

Cubeful Equities:
       No double:     +0,810 (-0,153)
       Double/Take:   +0,963
       Double/Pass:   +1,000 (+0,037)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   66,87% (G:24,43% B:1,77%)
  Opponent Winning Chances: 33,13% (G:9,69% B:0,23%)
Double/Take
  Player Winning Chances:   66,90% (G:25,01% B:1,87%)
  Opponent Winning Chances: 33,10% (G:9,86% B:0,24%)

Cubeless Equities: No Double=+0,500, Double=+1,012

Cubeful Equities:
       No double:     +0,728 (-0,042)
       Double/Take:   +0,770
       Double/Pass:   +1,000 (+0,230)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 19

```{r}
ggboard("XGID=--B-B-DBA---gB---c-dB---a-:0:0:1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   73,91% (G:5,95% B:0,08%)
  Opponent Winning Chances: 26,09% (G:1,39% B:0,06%)
Double/Take
  Player Winning Chances:   74,50% (G:6,10% B:0,12%)
  Opponent Winning Chances: 25,50% (G:1,28% B:0,07%)

Cubeless Equities: No Double=+0,548, Double=+1,356

Cubeful Equities:
       No double:     +0,873 (-0,127)
       Double/Take:   +1,073 (+0,073)
       Double/Pass:   +1,000

Best Cube action: Double / Pass

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   74,32% (G:5,85% B:0,08%)
  Opponent Winning Chances: 25,68% (G:1,38% B:0,04%)
Double/Take
  Player Winning Chances:   74,68% (G:5,67% B:0,08%)
  Opponent Winning Chances: 25,32% (G:1,36% B:0,04%)

Cubeless Equities: No Double=+0,532, Double=+1,074

Cubeful Equities:
       No double:     +0,820 (-0,075)
       Double/Take:   +0,894
       Double/Pass:   +1,000 (+0,106)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 20

```{r}
ggboard("XGID=----A-EbC---dD---b-cbAA-b-:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   66,52% (G:29,61% B:1,02%)
  Opponent Winning Chances: 33,48% (G:7,47% B:0,42%)
Double/Take
  Player Winning Chances:   66,45% (G:30,52% B:1,42%)
  Opponent Winning Chances: 33,55% (G:7,93% B:0,52%)

Cubeless Equities: No Double=+0,682, Double=+1,395

Cubeful Equities:
       No double:     +0,849 (-0,151)
       Double/Take:   +1,085 (+0,085)
       Double/Pass:   +1,000

Best Cube action: Double / Pass

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   66,50% (G:28,75% B:0,63%)
  Opponent Winning Chances: 33,50% (G:7,23% B:0,29%)
Double/Take
  Player Winning Chances:   66,58% (G:30,08% B:0,64%)
  Opponent Winning Chances: 33,42% (G:7,36% B:0,30%)

Cubeless Equities: No Double=+0,549, Double=+1,125

Cubeful Equities:
       No double:     +0,729 (-0,151)
       Double/Take:   +0,879
       Double/Pass:   +1,000 (+0,121)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

### Position 21

```{r}
ggboard("XGID=-B-B-bC-B--BbB--ac-cAAbb--:0:0:-1:00:0:0:0:0:10")
```

```{r}
#| layout-ncol: 2
"3-away, 3-away
Analyzed in Rollout
No double
  Player Winning Chances:   61,69% (G:28,96% B:0,79%)
  Opponent Winning Chances: 38,31% (G:9,37% B:0,49%)
Double/Take
  Player Winning Chances:   61,99% (G:29,19% B:1,65%)
  Opponent Winning Chances: 38,01% (G:11,08% B:0,61%)

Cubeless Equities: No Double=+0,539, Double=+1,058

Cubeful Equities:
       No double:     +0,588 (-0,152)
       Double/Take:   +0,741
       Double/Pass:   +1,000 (+0,259)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()

"Moneygame
Analyzed in Rollout
No double
  Player Winning Chances:   61,56% (G:28,46% B:0,72%)
  Opponent Winning Chances: 38,44% (G:8,96% B:0,30%)
Double/Take
  Player Winning Chances:   61,93% (G:28,83% B:0,67%)
  Opponent Winning Chances: 38,07% (G:8,85% B:0,32%)

Cubeless Equities: No Double=+0,430, Double=+0,884

Cubeful Equities:
       No double:     +0,517 (-0,085)
       Double/Take:   +0,602
       Double/Pass:   +1,000 (+0,398)

Best Cube action: Double / Take

Rollout:
  1296 Games rolled with Variance Reduction." %>% 
  remove_stuff() %>% 
  cat()
```

```{r show}
#| results: asis
#| include: false
#| eval: false


# Print all positions that we found
for (i in 1:nrow(df)) {
  cat("\n## Position ", i)
  ggboard(df$xgid[i]) %>% print()
  
  evaluation <- df$cube_eq[i] %>% 
    read_lines() %>% 
    paste(rep("   ", length(.)), .) %>% 
    paste(collapse = "\n")
  
  cat("\n\n", evaluation, 
      "\n\n    ", df$xgid[i], 
      "\n\n", sep = "")
}
```

```{r save}
#| include: false
#| eval: false

# Save the moneygame variant of all positions 
df$xgid %>% 
  str_split(":") %>% 
  map(~ replace(.x, 5, "00")) %>%  # Turn cube decision
  map(~ replace(.x, 9, "0")) %>%   # Turn into money game
  map_chr(~ paste(.x, collapse = ":")) %>% 
  write_lines(file = "data/good-cube-ids.txt")
```
