---
title: "Milestones III, 2023"
author: "Lasse Hjorth Madsen"
date: today 
format: 
  html:
    fig-width: 4
    fig-height: 4
toc: true
toc-expand: true
editor: source
execute:
  echo: false
  warning: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(kableExtra)

options(scipen = 9999)
devtools::load_all(path = "../../backgammon")

met_file <- devtools::package_file("data-raw/Kazaross XG2.met", path = "../../backgammon")
met <- get_met(filename = met_file)
```

## Why this?

This is a collection of positions from the Milestones III tournament in Denmark, 2023, where players compete for wins and for best performance rating (PR). Every player gets to play 15 7-point matches. I will collect about one problem from each of my own matches: Positions where I made some mistake, that I found difficult, surprising or otherwise interesting.

In some cases, I show also variations, to better understand what is going on. I will organize the positions only into cube actions and checker plays, adding more as they come.

## Cube actions

### Position 1

```{r}
ggboard("XGID=bA-a-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:2:0:0:7:10")
```

Over the board, this looked to me as a strong double. White has an excellent forward structure, a fat racing lead, and a four prime that can be extended from the back in relative safety, since Black has two checkers on the bar. The only weakness is White's blot on the ace point, but getting hit is not a big deal, so it can't be much of a liability. (Might it even be desirable to get hit, to have that checker recirculated back into action?)

Meanwhile, Black has made almost no progress; only the 11-point is made.

My guess was a strong double, and a smallish take for Black. It is, however, the other way around: White has only a small double, Black has a huge take:

```{r}
"Cubeful Equities:
       No double:     +0,690 (-0,047)
       Double/Take:   +0,736
       Double/Pass:   +1,000 (+0,264)
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller" %>%
cat()
```
Looking at the outcome probabilities, we see that White does have a big gammon threat, but also that Black wins more than a third of the games:

```{r}
# Analyzed in Rollout
# No double
#   Player Winning Chances:   65,12% (G:35,26% B:1,62%)
#   Opponent Winning Chances: 34,88% (G:7,79% B:0,49%)
# Double/Take
#   Player Winning Chances:   65,58% (G:36,41% B:2,44%)
#   Opponent Winning Chances: 34,42% (G:7,92% B:0,75%)

probs <- outcome_probs(c(65.58, 36.41, 2.44, 34.42, 7.92, 0.75)) %>% 
  `/`(100) %>% 
  round(3)

probs_fliped <- c(probs[4:6], probs[1:3])

probs %>% 
  probs_table() %>% 
  kable("html") %>%
  column_spec(2:5, width = "3.1cm")
```

One thing that I might have overlooked here, is that White has very few real crushing sequencies, because he's a little short on ammo. Black will almost always make some kind of anchor, and then have some kind of counterplay agains White's three back checkers.

To see if this result is match-score related, let's check the same position for money, and while we're at it, also check if XG Roller ++ (fast, truncated rollouts), gets it right, compared to a full rollout at strong settings. 

```{r}
#| layout-ncol: 2

ggboard("XGID=bA-a-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Cubeful Equities:
       No double:     +0,800 (-0,120)
       Double/Take:   +0,920
       Double/Pass:   +1,000 (+0,080)
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller

Analyzed in XG Roller++
       No double:     +0,784 (-0,134)
       Double/Take:   +0,918
       Double/Pass:   +1,000 (+0,082)
" %>%
cat("\n", ., sep = "") 
```

For money, the double is bigger, and the take is smaller, so more in line with my first impression. It appears that the 2-point lead for White does make a difference.

But why?

The next table shows Black's takepoints under different assumption: No cube and no gammons (like a last-roll position); no cube and gammon rates from the table above; and finally taking both cube and gammons into account. 

```{r}
tp_table <- tibble(
  `Cube assumptions` = c("Dead cube, no gammons", "Dead cube, gammons", "Both cube and gammons"),
  `Moneygame takepoint` = c(tp_money(c(0.5, 0, 0, 0.5, 0, 0), 0),
                            tp_money(probs_fliped, x = 0),
                            tp_money(probs_fliped, x = 0.68)
                            ),
  `7-away, 5-away takepoint` = c(tp(7, 5, 1, met, last_roll = TRUE),
                                 tp_info(7, 5, probs_fliped, 1, met)["tp_dead"],
                                 tp_info(7, 5, probs_fliped, 1, met)["tp_real"])
  )

tp_table %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `7-away, 5-away takepoint` - `Moneygame takepoint`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

The last-roll takepoint is not to different from money at this score, and it also doesn't make much difference to factor in gammons. When we consider possible recubes, however, we see that Black's takepoint at the score, is somewhat lower than it would be for money.

This is because he'll redouble more aggressively at the score. We can see why, by looking at *White's* takepoints for a potential recube, assuming the same gammon frequencies: 

```{r}
tp_table <- tibble(
  `Cube assumptions` = c("Dead cube, no gammons", "Dead cube, gammons", "Both cube and gammons"),
  `Moneygame takepoint` = c(tp_money(c(0.5, 0, 0, 0.5, 0, 0), 0),
                            tp_money(probs, x = 0),
                            tp_money(probs, x = 0.68)
                            ),
  `5-away, 7-away recube takepoint` = c(tp(5, 7, 2, met, last_roll = TRUE),
                                 tp_info(5, 7, probs, 2, met)["tp_dead"],
                                 tp_info(5, 7, probs, 2, met)["tp_real"])
  )

tp_table %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `5-away, 7-away recube takepoint` - `Moneygame takepoint`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```


## Checker plays

