---
title: "Milestones III, 2023"
author: "Lasse Hjorth Madsen"
date: today 
format: html
toc: true
toc-expand: true
editor: source
execute:
  echo: false
  warning: false
---
<style type="text/css">
   .main-container {max-width: 100%;}
   .row {display: flex;}
   .column {flex: 100%;}
</style>

```{r setup}
library(tidyverse)
library(knitr)
library(kableExtra)
library(DT)
options(scipen = 9999)
options(knitr.table.format = "html")
devtools::load_all(path = "../../backgammon")
```

## Why this?

This is a collection of positions from the Milestones III tournament in Denmark, 2023, where players compete for wins and for best performance rating (PR). Every player gets to play 15 7-point matches. I will try to collect one problem from each of my own matches: Positions where I made some mistake, that I found difficult, surprising or otherwise interesting.

In some cases, I show also variations, to better understand what is going on. I will try to organize the positions in some broad categories.

## Cube actions

<div class = "row">
<div class = "column">
```{r}
#| fig.width: 4
#| fig.height: 4
ggboard("XGID=bA-a-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:2:0:0:7:10")
```
</div>


```{r}
"Cubeful Equities:
       No double:     +0,690 (-0,020)
       Double/Take:   +0,710
       Double/Pass:   +1,000 (+0,290)
Best Cube action: Double / Take" %>%
  read_lines() %>% 
  as_tibble() %>% 
  slice(c(-1,-5)) %>% 
  separate(value, into = c("Action", "Equity"), sep = ":") %>%
  datatable(options = list(dom = 't', ordering = F, headerCallback = DT::JS(
    "function(thead) {",
    "  $(thead).css('font-size', '80%');",
    "}"
  )), width = "100%", rownames = NULL) %>% 
  formatStyle(columns = c(1, 2), fontSize = '80%')
```
</div>

## Checker plays

