---
title: "Milestones III, 2023"
author: "Lasse Hjorth Madsen"
date: today 
format: 
  html:
    fig-width: 4
    fig-height: 4
toc: true
toc-expand: true
editor: source
execute:
  echo: false
  warning: false
  cache: true
---

```{r setup}
#| cache: false
library(tidyverse)
library(knitr)
library(kableExtra)

options(scipen = 9999)
devtools::load_all(path = "../../backgammon")

met_file <- devtools::package_file("data-raw/Kazaross XG2.met", path = "../../backgammon")
met <- get_met(filename = met_file)
```

## Why this?

This is a collection of positions from the Milestones III tournament in Denmark, 2023, where players compete for wins and for best performance rating (PR). Every player gets to play 15 7-point matches. I will collect about one problem from each of my own matches: Positions where I made some mistake, that I found difficult, surprising or otherwise interesting.

In some cases, I show also variations, to better understand what is going on. I will organize the positions only into cube actions and checker plays, adding more as they come.

## Cube actions

### Position 1

```{r}
ggboard("XGID=bA-a-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:2:0:0:7:10")
```

Over the board, this looked to White like a strong double. White has an excellent forward structure, a fat racing lead, and a 4-point prime that can be extended from the back in relative safety, since Black has two checkers on the bar. The only weakness is White's blot on the ace point, but getting hit is not a big deal, so it can't be much of a liability.

Meanwhile, Black has made almost no progress; only the 11-point is made.

White guessed this was a strong double, and a smallish take for Black. It is, however, the other way around: White has only a small double, Black has a huge take:

```{r}
"Cubeful Equities:
       No double:     +0,690 (-0,047)
       Double/Take:   +0,736
       Double/Pass:   +1,000 (+0,264)
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller" %>%
cat()
```

Looking at the outcome probabilities, we see that White does have a big gammon threat, but also that Black wins more than a third of the games:

```{r}
# Analyzed in Rollout
# No double
#   Player Winning Chances:   65,12% (G:35,26% B:1,62%)
#   Opponent Winning Chances: 34,88% (G:7,79% B:0,49%)
# Double/Take
#   Player Winning Chances:   65,58% (G:36,41% B:2,44%)
#   Opponent Winning Chances: 34,42% (G:7,92% B:0,75%)

probs <- outcome_probs(c(65.58, 36.41, 2.44, 34.42, 7.92, 0.75)) %>% 
  `/`(100) %>% 
  round(3)

probs_fliped <- c(probs[4:6], probs[1:3])

probs %>% 
  probs_table() %>% 
  kable("html") %>%
  column_spec(2:5, width = "3.1cm")
```

One thing that might be overlooked here, is that White has very few real crushing frequencies, because he's a little short on ammo. Black will almost always make some kind of anchor, and then have some kind of counter play agains White's three back checkers.

To see if this result is match-score related, let's check the same position for money, and while we're at it, also check if XG Roller ++ (fast, truncated rollouts), gets it right, compared to a full rollout at strong settings.

```{r}
#| layout-ncol: 2

ggboard("XGID=bA-a-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Cubeful Equities:
No double:     +0,800 (-0,120)
Double/Take:   +0,920
Double/Pass:   +1,000 (+0,080)
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller

Analyzed in XG Roller++
No double:     +0,784 (-0,134)
Double/Take:   +0,918
Double/Pass:   +1,000 (+0,082)
" %>%
cat("\n", ., sep = "") 
```

For money, the double is bigger, and the take is smaller, so more in line with White's first impression. It appears that the 2-point lead for White does make a difference.

But why?

The next table shows Black's take points under different assumption: a) No cube and no gammons (like a last-roll position); b) no cube and gammon rates from the table above; and c) taking both cube and gammons into account.

```{r}
tp_table <- tibble(
  `Cube assumptions` = c("Dead cube, no gammons", "Dead cube, gammons", "Both cube and gammons"),
  `money game take point` = c(tp_money(c(0.5, 0, 0, 0.5, 0, 0), 0),
                            tp_money(probs_fliped, x = 0),
                            tp_money(probs_fliped, x = 0.68)
                            ),
  `7-away, 5-away take point` = c(tp(7, 5, 1, met, last_roll = TRUE),
                                 tp_info(7, 5, probs_fliped, 1, met)["tp_dead"],
                                 tp_info(7, 5, probs_fliped, 1, met)["tp_real"])
  )

tp_table %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `7-away, 5-away take point` - `money game take point`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

The last-roll take point is not too different from money at this score, and it also doesn't make much difference to factor in gammons. When we consider possible recubes, however, we see that Black's take point at the score, is somewhat lower than it would be for money.

This is because he'll redouble more aggressively at the score. We can see why, by looking at *White's* take points for a potential recube, assuming the same gammon frequencies:

```{r}
tp_table <- tibble(
  `Cube assumptions` = c("Dead cube, no gammons", "Dead cube, gammons", "Both cube and gammons"),
  `money game take point` = c(tp_money(c(0.5, 0, 0, 0.5, 0, 0), 0),
                            tp_money(probs, x = 0),
                            tp_money(probs, x = 0.68)
                            ),
  `5-away, 7-away recube take point` = c(tp(5, 7, 2, met, last_roll = TRUE),
                                 tp_info(5, 7, probs, 2, met)["tp_dead"],
                                 tp_info(5, 7, probs, 2, met)["tp_real"])
  )

tp_table %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `5-away, 7-away recube take point` - `money game take point`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

So, it looks like a good part of White's flawed evaluation of the original position, comes from the fact that Black can take somewhat more aggressively at the score; hence White needs more to have a really strong double.

The reason for Black's lower take point at the score, lies mostly in the potential redouble in positions where gammons are possible: It's hard for White to accept a 4-cube at 5-away 7-away, where a gammon loss is much more costly than a gammon win.

Still, the position is also a clear take for money, so there's some reconsideration for White to do in that department too. What will it take to turn the position into a really tough take for money?

If we activate the awkward spare on the ace-point, we suddenly have a big pass. Perhaps not surprising, since we're turning a big liability into a big asset. But still:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BCBB-A-cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,979 (-0,021)
Double/Take:   +1,273 (+0,273)
Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

Even if we put the spare checker in a less ideal place, we still get a double/pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BBBB-A-cCb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,929 (-0,071)
Double/Take:   +1,045 (+0,045)
Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

That spare has to move out of the action, to get to a marginal pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BBBB-A-cBb-Ac-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,913 (-0,087)
Double/Take:   +1,011 (+0,011)
Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

Finally, we get to a decent take, noting that Black is no longer far behind in the race:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BBBB-A-cBb--c-dB-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,874 (-0,072)
Double/Take:   +0,947
Double/Pass:   +1,000 (+0,053)" %>%
cat("\n", ., sep = "") 
```

One way to see how bad White's ace-point blot is, is to consider the original position after something like 31: 24/20. Now, most aces, like 21, 31, 41, and 51, are below-average rolls for Black, even though they hit. White is *hoping* to get hit (except with 41 and 11, trying for a high anchor); that's how bad the blot is.

Another experiment would be to improve White's opportunity to extend the prime from the back. A one-pip change can be huge: We go from a clear take to a pretty clear pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=bA-a-BBBBA--cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,856 (-0,144)
Double/Take:   +1,059 (+0,059)
Double/Pass:   +1,000
" %>%
cat("\n", ., sep = "") 
```

Another change in the same spirit is to bring in another builder for that important 9-point. It has about the same effect as slotting the 9-point:

```{r}
#| layout-ncol: 2
ggboard("XGID=bA-a-BBBB-A-cBbA-c-d--A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,854 (-0,146)
Double/Take:   +1,065 (+0,065)
Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

A final one-pip change that has a rather dramatic effect, is to just semi-active the ace-point checker so it easier becomes part of a prime, and is both easier and more attractive to cover. Again, we go from a clear take to a clear pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=b-Aa-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,849 (-0,151)
Double/Take:   +1,062 (+0,062)
Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

Summing up:

-   A two-point lead in a 7-point match matters, especially if you might face a recube when gammons are still possible.
-   What seems to be a strong positional advantage, might not be so strong if there are also liabilities.
-   An awkward ace-point blot can be a rather big liability
-   Small things matter: One pip can make a it much easier to turn a strong position into a winning position.

### Position 2

```{r}
ggboard("XGID=bAB--BBAB---bB--Abbcc-AAa-:0:0:1:00:2:0:0:7:10")
```
This is another example of blitz that could work as a bluff double: It might look like White has a strong blitz going, but in fact the take is easy, and this is not even a double. In practice, it's not impossible to get a pass. (In the actual game, Black passed this.)

```{r}
"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,579
Double/Take:   +0,534 (-0,045)
Double/Pass:   +1,000 (+0,421)" %>% 
  cat()
```

At the match score, the error from passing is gigantic, for money it is smaller but still big:

```{r}
#| layout-ncol: 2
ggboard("XGID=bAB--BBAB---bB--Abbcc-AAa-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,664 (-0,078)
Double/Take:   +0,742
Double/Pass:   +1,000 (+0,258)" %>% 
  cat("\n", ., sep = "")
```

To better understand why this score is so different from money, we can compare the take points for money and at the score, using different assumptions:

```{r}
# Analyzed in XG Roller++
# Player Winning Chances:   67,99% (G:36,77% B:1,15%)
# Opponent Winning Chances: 32,01% (G:9,36% B:0,46%)

probs <- outcome_probs(c(67.99, 36.77, 1.15, 32.01,9.36, 0.46)) %>% 
  `/`(100) 

probs_fliped <- c(probs[4:6], probs[1:3])
```

```{r}
tp_tab <- tp_table(7, 5, probs_fliped, 1, met)

tp_tab %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `7-away, 5-away take point` - `money game take point`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

As in position 1, it is the powerful recube access that makes Black considerable more inclined to take at the score, compare to money game. 

Let's keep the money game variant, and play with features of the position. 

The ace-point slots makes all the difference in the world, because it commits White to a blitz. If we turn the slot into a builder White can both blitz or prime, and two strong game plans are much better than one. As a result Black has a moderate sized pass in the variant below. 

```{r}
#| layout-ncol: 2
ggboard("XGID=b-B--BCAB---bB--Abbcc-AAa-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,865 (-0,135)
Double/Take:   +1,067 (+0,067)
Double/Pass:   +1,000" %>% 
  cat("\n", ., sep = "")
```

Also, making it a bit easier for White to jump his back checkers, turns the position into just a borderline take.

```{r}
#| layout-ncol: 2
ggboard("XGID=bAB--BBAB---bB--AbbccAA-a-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,784 (-0,211)
Double/Take:   +0,996
Double/Pass:   +1,000 (+0,004)" %>% 
  cat("\n", ., sep = "")
```

In summary, it is the combination of having to jump a prime, without being on the edge, *and* being committed to a blitz, *and* being a little short on ammunition, that makes Whites's position weaker than it might appear.

### Position 3

```{r}
ggboard("XGID=abB---CAB--AcCAa-c-e----B-:0:0:1:00:4:2:0:7:10")
```

White had a good start, and considered doubling, but must be mindful of the score, leading 3-away, 5-away. Is the position strong enough to double?

No, doubling would be a big mistake:

```{r}
"Analyzed in XG Roller++
No double:     +0,452
Double/Take:   +0,160 (-0,292)
Double/Pass:   +1,000 (+0,548)" %>% 
  cat()
```

To better understand the position, consider two illustrative variants: Money game (left), and money game with White's 2-point switched back a pip (right) so it becomes the 3-point:

```{r}
#| layout-ncol: 2
ggboard("XGID=abB---CAB--AcCAa-c-e----B-:0:0:1:00:0:0:0:0:10")
ggboard("XGID=ab-B--CAB--AcCAa-c-e----B-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,619
Double/Take:   +0,556 (-0,063)
Double/Pass:   +1,000 (+0,381)" %>% 
  cat()

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,731 (-0,011)
Double/Take:   +0,742
Double/Pass:   +1,000 (+0,258)" %>% 
  cat()
```

The original is still a no-double for money, although doubling is a much smaller mistake that at the score. Shifting to a better inner point makes a clear difference; now the decision to double is really close, and White has a small double. 

Bringing in more ammunition *and* making the 11-point is required to turn the position into a solid double/take for money, although it would still be a clear no-double at the score:

```{r}
#| layout-ncol: 2
ggboard("XGID=ab-B--CAC--BcB-a-c-e----B-:0:0:1:00:4:2:0:7:10")
ggboard("XGID=ab-B--CAC--BcB-a-c-e----B-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,686
Double/Take:   +0,581 (-0,105)
Double/Pass:   +1,000 (+0,314)" %>% 
  cat()

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,838 (-0,082)
Double/Take:   +0,920
Double/Pass:   +1,000 (+0,080)" %>% 
  cat()
```

Only if we in addition to the modifications above, *also* switch back to the much stronger 4-point, do we have a double at the score, but only just. For money, the position is now a big pass.

```{r}
#| layout-ncol: 2
ggboard("XGID=ab--B-CAC--BcB-a-c-e----B-:0:0:1:00:4:2:0:7:10")
ggboard("XGID=ab--B-CAC--BcB-a-c-e----B-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,847 (-0,011)
Double/Take:   +0,859
Double/Pass:   +1,000 (+0,141)" %>% 
  cat()

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,939 (-0,061)
Double/Take:   +1,164 (+0,164)
Double/Pass:   +1,000" %>% 
  cat()
```

Let's compare take points at the score and for money (using outcome probabilities from the original position) to see why cube action is so different at this score than for money:

```{r}
# Analyzed in XG Roller++
# Player Winning Chances:   61,53% (G:26,11% B:3,01%)
# Opponent Winning Chances: 38,47% (G:9,32% B:0,52%)

probs <- outcome_probs(c(61.53, 26.11, 3.91, 38.47, 9.32, 0.52)) 
probs_fliped <- c(probs[4:6], probs[1:3])

tp_tab <- tp_table(5, 3, probs_fliped, 1, met)

tp_tab %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `5-away, 3-away take point` - `money game take point`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

While it would actually be harder for Black to accept a double in a short race (a 4-roll position would be a pass), the ability to redouble makes all the difference.

The take points are based on the below outcome probabilities from XG roller++. Note that Black has as much as 38 percent to win this game (in the original position):

```{r}
probs %>% 
  probs_table() %>% 
  kable("html") %>%
  column_spec(2:5, width = "3.1cm")
```

In summary, when leading 3-away, 5-away, in a typical middle game position, the leader should have a position that is a clear money pass, to consider doubling. That potential re-cube is just too powerful to let your opponent have.

### Position 4
```{r}
ggboard("XGID=-aa--BBBC--BAB-b-bbe-bA---:0:0:1:00:4:0:0:7:10")
```

Here, White figured that escaping with a 6 or making the 9- or 10-point, could lead to a market loss and doubled. White failed to pay enough attention to the match score, *and* overestimated the position a bit, the combination of which made the decision to double a 0.308 blunder.

The three diagrams below show left to right: a) the position for money: A borderline double at best; b) The modification required to turn it into a solid double and a solid take; c) The closest variant that would make the position only a small take for money 

```{r}
#| layout-ncol: 3
ggboard("XGID=-aa--BBBC--BAB-b-bbe-bA---:0:0:1:00:0:0:0:0:10")
ggboard("XGID=-aa--BBBC-B-AB-b-bbe-bA---:0:0:1:00:0:0:0:0:10")
ggboard("XGID=-aa--BBBC-B-AB-b-bbeAb----:0:0:1:00:0:0:0:0:10")
```

The market loss that White imagined was something like a 63: 22/13, which would indeed have lost White's market -- if Black was not allowed to roll. But Black gets to roll, and most rolls make some kind of improvement that allows Black to take, so true market losers are rare.

But the big part of the mistake was failing to account for the match score. It's easy to see how the ability to recube to 4 is a powerful thing for Black when trailing 7-away, 3-away. After a redouble/take, Black can win 4 points, with the added possibility of winning a gammon and the match; White can win only 3 points and can't benefit from gammons.

Let's line up Black's take points in a table and compare to money game:

```{r}
# Analyzed in XG Roller++
# Player Winning Chances:   69,98% (G:13,65% B:0,65%)
# Opponent Winning Chances: 30,02% (G:7,14% B:0,23%)

probs <- outcome_probs(c(69.98, 13.65, 0.65, 30.02, 7.14, 0.23)) 
probs_fliped <- c(probs[4:6], probs[1:3])

tp_tab <- tp_table(7, 3, probs_fliped, 1, met)

tp_tab %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `7-away, 3-away take point` - `money game take point`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

Like we saw in an earlier problem, Black can take really deep at this score and these gammon frequencies -- take point all things considered is about 19%.

For money White can justify a cube on a sunny day; at the match score he needs a better position.

The variant that makes for a marginal money game take, for example, would still be a sizable no-double at the score:

```{r}
#| layout-ncol: 2

ggboard("XGID=-aa--BBBC-B-AB-b-bbeAb----:0:0:1:00:4:0:0:7:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:   +0,731
Double/Take: +0,573 (-0,158)
Double/Pass: +1,000 (+0,269)" %>%
  cat("\n", ., sep = "") 
```

### Position 5

```{r}
ggboard("XGID=aB-b-CB-C---bD--bb-ccA----:0:0:1:00:2:0:0:7:10")
```

This one is much the same theme as the previous position: White overestimated their winning chances and figured that escaping the back checker would be a clear market loss. In fact, this is a big no double:

```{r}
"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,385
Double/Take:   +0,114 (-0,271)
Double/Pass:   +1,000 (+0,615)" %>% 
  cat()
```

The mistake is magnified by the score, but for money it is still a massive no-double:

```{r}
#| layout-ncol: 2
ggboard("XGID=aB-b-CB-C---bD--bb-ccA----:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,488
Double/Take:   +0,343 (-0,145)
Double/Pass:   +1,000 (+0,512)" %>% 
  cat("\n", ., sep = "") 
```

It is easy to see the mistake considering that escaping *and* having Black fan, will not even turn it into a double, much less a drop:

```{r}
#| layout-ncol: 2
ggboard("XGID=aB-b-CB-C---bE--bb-cc-----:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,747
Double/Take:   +0,713 (-0,035)
Double/Pass:   +1,000 (+0,253)" %>%
  cat("\n", ., sep = "")
```

In fact, to make this position a close take/pass decision, White needs pretty massive improvements, like this:

```{r}
#| layout-ncol: 2
ggboard("XGID=aB-b-BCABB--bC--bb-cc-----:0:0:1:00:0:0:0:0:10")

"Analyzed in Rollout
Cubeful Equities:
No double:     +0,913 (-0,076)
Double/Take:   +0,989
Double/Pass:   +1,000 (+0,011)" %>% 
  cat("\n", ., sep = "") 
```

The weakness of White's position comes in large part from the the gap on the 4-point. If we close that (by sliding Black's anchor forward one pip), the the escape/fan-sequence is now a double/take. (The original position, of course, remains a no-double, even with the gap closed, since escape/fan is still not a market loss).

```{r}
#| layout-ncol: 2
ggboard("XGID=aB--bCB-C---bE--bb-cc-----:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,877 (-0,043)
Double/Take:   +0,920
Double/Pass:   +1,000 (+0,080)" %>% 
  cat("\n", ., sep = "") 
```

### Position 6

```{r}
ggboard("XGID=-B----C-D-a-eCA--c-dbA--A-:0:0:-1:00:4:0:0:7:10")
```

In this position White underestimated an attacking position, *and* failed to take the match score into account, and so took a double that is a big pass: 

```{r}
"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,984 (-0,016)
Double/Take:   +1,300 (+0,300)
Double/Pass:   +1,000" %>% 
  cat()
```

The fact that White can lose a costly gammon and can almost never redouble, is what makes the error really big. Black's gammon value after double/take is `r gammon_value(7, 3, 2, met) %>% round(2)`; White's gammon value is `r gammon_value(3, 7, 2, met) %>% round(2)`. Should White decide to redouble, Black can pass and get to match winning chance of `r mwc(7,1, met) %>% round(2)` or take and redouble if he is doing any better than that. So White can't get much benefit from the cube either.

But the position itself is also quite bad, a clear money pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=-B----C-D-a-eCA--c-dbA--A-:0:0:-1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,914 (-0,086)
Double/Take:   +1,089 (+0,089)
Double/Pass:   +1,000" %>% 
  cat("\n", ., sep = "") 
```

Part of the problem with White's position is the made ace-point; sliding that point back to the 3-point results in a marginal take/pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=---B--C-D-a-eCA--c-dbA--A-:0:0:-1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,884 (-0,112)
Double/Take:   +0,996
Double/Pass:   +1,000 (+0,004)" %>% 
  cat("\n", ., sep = "") 
```

Of course, removing the annoying double shot that Black has in the outfield, also improves White's position a lot. Clear money-take here (although still a pass at the original score):

```{r}
#| layout-ncol: 2
ggboard("XGID=-B----C-D-a-eD---c-dbA--A-:0:0:-1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
No double:     +0,810 (-0,069)
Double/Take:   +0,879
Double/Pass:   +1,000 (+0,121)
" %>% 
  cat("\n", ., sep = "") 
```

To turn the position into a take at 3-away, 7-away, White needs a series of improvements: A big upgrade of the ace-point into the 5-point, removing the 14-point shot, *and* placing Black's escaped checker a little further from home:

```{r}
#| layout-ncol: 2
ggboard("XGID=-----BC-Da--eD---c-dbA--A-:0:0:-1:00:4:0:0:7:10")

"Analyzed in Rollout
Cubeful Equities:
No double:     +0,892 (-0,106)
Double/Take:   +0,997
Double/Pass:   +1,000 (+0,003)" %>% 
  cat("\n", ., sep = "") 
```

Black is still a lot better off than White since he has nearly escaped both back checkers, and is well ahead in the race. Black wins about 70% of the games, but White can probably just squeeze out a very marginal take at the score.

## Checker plays

### Position 1

```{r}
ggboard("XGID=-b--B-CBB--BcB-b---db-b-B-:0:0:1:52:2:1:0:7:10")
```

Here, White has four reasonable options:

- 13/11, 13/8, clearing the midpoint
- 11/4, placing a spare on the 4-point 
- 11/6, 11/9, with spares on the 6- and 9-points 

Unwilling to break the midpoint as long as reasonable alternatives existed, White played 11/4, wanting to leave fewer shots and cover more outfield, compared to 11/6, 11/9.

It is not hard to accept that this can be a mistake:

11/6, 11/9 also breaks the 11-point, but is more compact. It adds 63 to build the 3-point, slots the 9-point which could be useful, and only leaves 44 as an extra shot (a joker anyway).

13/11, 13/8, clearing the midpoint, puts full emphasis on White's all-important 5-point, leaves no shot, and makes it more dangerous for Black to split with a roll like 42. 
 
The interesting part is not that XG picks 13/11, 13/8, but the *magnitude* of the mistake is remarkable; anything but breaking the midpoint is a blunder: 

```{r}
"XG Roller++ 
1. 13/11 13/8           eq:+0,026
2. 11/9 11/6            eq:-0,065 (-0,090)
3. 11/4                 eq:-0,131 (-0,157)" %>% 
  cat()
```

The key is, that after the top play, 9 rolls make a full 5-prime. After the second best play, only 4 rolls do that. Some rolls, 31 and 43, makes the 5-point but breaks the 8, which is not as strong. So the extra builders on the 11- and 8-points have real value. 

Note that White should break the 13-point with a lot of rolls: 54, 43, 62, 42, 32, 51, and the 52 from the actual position.

51 deserves special mention, the funny looking 13/7 is probably slightly better than the aggressive slot, 11/5, which accomplishes the same thing, but with a different risk/reward ratio: A lot more shots, and a lot more ways to make the 5-point if missed.

The shots left in the outfield is not the main concern: 43 should be played 13/10 13/9 with lots of builders at the cost of 5 shots; 43: 11/8, 11/7 would be a blunder even though it looks flexible.

Likewise, with 32, 13/11, 13/10 is a lot better than 11/9, 11/8, since it creates four active builders for the 5-point rather than just three.

Summing up, making the 5-point and a 5-point-prime is likely to win the game, and White can easily afford to break the midpoint and leave a few fly-shots to increase the probability of achieving that goal.

Builders win out -- *if* they aim at something important, and *if* they are not too likely to get hit.

### Positions 2a-2e

```{r}
#| layout-ncol: 2
ggboard("XGID=-a-BbCCBB-----c-Bcac-b-A--:1:1:1:42:3:0:0:7:10") + 
  labs(title = "Position 2a")
ggboard("XGID=-aB--CC-C---bC--abbdc---A-:0:0:1:51:2:0:0:7:10") + 
  labs(title = "Position 2b")
ggboard("XGID=-ABbBBB-C---cB---c-cbA--b-:1:-1:1:21:0:0:0:7:10") + 
  labs(title = "Position 2c")
ggboard("XGID=aABaB-C-CA--cB---c-cbAa-a-:1:-1:1:41:0:0:0:7:10") + 
  labs(title = "Position 2d")
ggboard("XGID=aAaBBBB-B---bB-aAa-cbbb-A-:1:-1:1:11:0:0:0:7:10") + 
  labs(title = "Position 2e")
```
These five positions share a common theme: The right play is hard to find for those of us who learned, back in the day, that the one thing you must never do in backgammon, is to make the ace-point.


```{r}
#| layout-ncol: 2

"Position 2a
XG Roller++ 
1. 5/1* 3/1     eq:+0,266
2. 8/6 5/1*     eq:+0,215 (-0,051)
3. 7/1*         eq:+0,204 (-0,062)" %>% 
  cat()

"Position 2b
XG Roller++ 
1. 6/1* 2/1     eq:-0,324
2. 8/7 6/1*     eq:-0,361 (-0,037)
3. 24/23 6/1*   eq:-0,413 (-0,089)" %>% 
  cat()

"Position 2c
XG Roller++ 
1. 8/6 2/1      eq:+0,178
2. 8/5          eq:+0,130 (-0,048)" %>% 
  cat()

"Position 2d
XG Roller++ 
1. 6/1          eq:+0,767
2. 8/3*         eq:+0,754 (-0,013)
3. 9/5 6/5      eq:+0,723 (-0,044)" %>% 
  cat()

"Position 2e
XG Roller++ 
1. 24/23 16/15* 3/2* 2/1    eq:+0,909
2. 24/23 16/15* 3/2*(2)     eq:+0,857 (-0,052)
3. 16/15* 8/7 3/2*(2)       eq:+0,850 (-0,059)
4. 16/15* 8/7 3/2* 2/1      eq:+0,818 (-0,091)
5. 24/23 16/15* 8/7(2)      eq:+0,804 (-0,105)" %>% 
  cat()
```
In position 2a and 2b White must hit, he has the stronger board and while it is not good to get hit back, Black can also do some damage if left alone. After hitting on the ace point, anything but switching point just leave another blot for no particular gain.

In 2c White just overlooked switching -- an easy thing to do if you learned to play pre-neural nets. 

2d is just a case of a classic blitz: It doesn't matter too much which points you make; in fact the ace-point may be better to make, since the higher points are easier to cover.

2e is a little more tricky prime vs. prime kind of thing. The important part is to but three checkers on the bar -- then White might have time to recover even if hit back. It's also good to advance with 24/23. So after three mandatory aces: 24/23, 16/15\*, and 3/2\*, it is easier to see that 2/1 is the right final ace, since it leaves a direct cover for the slotted 3-point.

The positional approach that White tried in the actual game, 24/23 16/15* 8/7(2), is no good; Black can both anchor or hit back. 

### Position 3

```{r}
ggboard("XGID=----A-FAB-A--C-----Adcaba-:1:1:1:64:0:0:0:0:10")
```

This is a technical play, where it is tempting to assume that all reasonable plays are close. Actually, they are not so close.

Obviously the play is all about getting off the gammon. Over the board, White played 19/9, thinking that it hardly mattered, as long as he made crossovers and avoided wastage by playing past the 6-point.

The gammon race is close. The probabilities that White will lose a gammon after the five best plays, are listed below. 

```{r}
"1. 13/7 10/6    53.93% 
2. 19/13 10/6   55.07%
3. 19/15 13/7   56.63%
4. 19/9         57.74%
5. 13/9 13/7    58,07%
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller" %>% 
  cat()
```
The actual play, 19/9, increases the risk of getting gammoned, by almost 4 percentage points, so for money that would be about a 0.04 mistake, costing 4 percent of the cube value -- a significant mistake.

Why so much?

The main concept in getting-off-the-gammon problems is to avoid "wasting" pips, by playing past the 6-point, those pips don't help saving the gammon. Playing 10/6 with the 4 does just that, so must be right. Indeed, 10/6 is part of both top plays. 

This *is* quite important, especially when the gammon race is close.

After 10/6 with the 4,  13/7 is best with the 6, simply because it creates a smaller stack: Two checkers on the 7- and 13-point, is better than one and four checkers respectively.

### Position 4

```{r}
ggboard("XGID=-----DF-C----A------A--be-:2:-1:1:31:0:0:0:7:10")
```
This one is from the same game as position 3, but Black has rolled well, so White is now a big favorite to get gammoned. The 31 White has to play is a little funny, because the correct play seem to violate basic principles for getting off the gammon: Make crossovers; don't waste pips.

The right play is 8/7 8/5, wasting a pip and failing to make two crossovers.

A little reflection shows that White needs to cater to specific doubles. After any play White needs 7 or 8 crossovers in three rolls, so one of roll has to be a double to save the gammon. After 8/7 8/5 any double makes 4 crossover, which is not true for the other plays, so that is the winner.

It's true, that if White makes two crossovers now, with 13/12 8/5, he only needs to generate three crossovers from a future double. But 11 fails to do even that, so it loses a tiny bit there.

The other two-crossover play, played in the actual game, is 20/17 13/12 which also fails on 11, but plays worse on 55 and 44 as well.

The probabilities of losing a gammon after the 7 best plays are:

```{r}
"1. 8/7 8/5      89,63% 
2. 13/12 8/5    89,78% 
3. 20/19 8/5    90,42% 
4. 8/4          90,99% 
5. 20/16        91,03% 
6. 20/19 13/10  91,09% 
7. 20/17 13/12  91,39% 
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller" %>% 
  cat()
```

### Position 5

```{r}
ggboard("XGID=-b---DC-C--AbB----acd-c-B-:1:1:1:64:0:0:0:7:10")
```

This is a tough play. White has two reasonable but very different plays: Hitting with 24/18\*/14, or priming with 13/7 11/7. 

Hitting has big upsides and big downsides. Black hits back on fully 18 rolls, but also has some bad rolls, 4 dancing rolls, and an additional 8 bad sixes. Some hitting rolls, like 51, aren't too happy either. 

Priming could work, since White is behind in the race. It leaves Black a few immediate horror rolls, 55 and 44, but also some pretty strong hitting aces. 

In the end, priming comes out slightly better:

```{r}
"1. 13/7 11/7          eq:+0,417
2. 24/18* 18/14       eq:+0,401 (-0,016)
3. 24/18* 11/7        eq:+0,239 (-0,178)" %>% 
  cat()
```

Let's have a look at the money game variant, together with a very minimal one-pip modification:

```{r}
#| layout-ncol: 2
ggboard("XGID=-b---DC-C--AbB----acd-c-B-:1:1:1:64:0:0:0:0:10")
ggboard("XGID=-b---DC-C--AbB----adc-c-B-:1:1:1:46:0:0:0:0:10")

"XG Roller++ 
1. 13/7 11/7          eq:+0,413
2. 24/18* 18/14       eq:+0,405 (-0,008)" %>% 
  cat()

"XG Roller++ 
1. 24/18* 18/14       eq:+0,367
2. 13/7 11/7          eq:+0,343 (-0,024)" %>% 
  cat()
```

For money priming is still right, but now super close; the two plays are basically a photo finish. 

But why the (small) difference compare to 7-away, 7-away score? The reason is, that the hitting play loses a bit more gammons than the priming play. At the score, the gammon value with the cube at two is `r gammon_value(7, 7, 2, met) %>% round(2)`, compare to 0.5 for money, so gammons are slightly more costly. For that reason, the hit is a bit more attractive for money, making the two plays even closer. 

The real mystery, however, is the variation on the right, where one Black checker from the 5-point has moved back a pip to the 6-point. This seemingly very minor modification makes the hitting play correct by a small but a meaningful amount: 0.024.

The reason is, probably, that the change is not so small after all. Note that the equities in the second position are quite a bit lower, so White is significantly worse off. For that reason, there's a bit less to lose, when the hitting play backfires.

Also, and perhaps more importantly, in a priming battle Black can use the extra spare on the 6-point to play future 4s, since that number is blocked on the other side of the board. For example, imagine White makes the priming play in both variations, and Black then rolls 54, making the 2-point. Now, the position with two spares on the 5-point really looks a lot worse than the variant with a spare on each of the 5- and 6-point. (And it is, evaluations confirm.)

Bottom line is that Black is better set-up for priming in the right-hand position, so White should chose another gameplan. 

Small differences often are not so small.

Let's do one more exercise with this position, and look at how a few analogous problems evaluate:

```{r}
#| layout-ncol: 2
ggboard("XGID=-b---DC-C--AbB----acd-c-B-:1:1:1:16:0:0:0:0:10")

"XG Roller++ 
1. 13/7 8/7           eq:+0,394
2. 24/23 24/18*       eq:+0,367 (-0,027)
3. 24/18* 18/17       eq:+0,327 (-0,067)
4. 24/18* 11/10       eq:+0,314 (-0,080)" %>% 
  cat("\n", ., sep = "") 

ggboard("XGID=-b---DC-CA--bB----acd-c-B-:1:1:1:26:0:0:0:0:10")

"XG Roller++ 
1. 13/7 9/7           eq:+0,411
2. 24/18* 18/16       eq:+0,302 (-0,109)
3. 9/3 5/3            eq:+0,290 (-0,121)
" %>% 
  cat("\n", ., sep = "") 

ggboard("XGID=-b---DC-C-A-bB----acd-c-B-:1:1:1:36:0:0:0:0:10")

"XG Roller++ 
1. 24/21 24/18*       eq:+0,427
2. 13/7 10/7          eq:+0,413 (-0,015)" %>% 
  cat("\n", ., sep = "") 
```

With a 61 to play, rather that a 64, priming performs better. Not so much because priming is better with 61, but because the alternative hitting plays are all worse than the hit-and-run-almost-home play available with 64.

In the second variation, with the same kind of choice for a roll of 62, priming is now a stand-out. Both because hitting and running still doesn't run as far, *and* (I think more importantly) because hitting leaves an additional 4 very dangerous returns from the bar: 63 and 54.

In the final variation, hitting is slightly right, likely because the 10-point would be valuable to make also, so that checker works also as a slot. (Which may be covered with a 5 from the 15-point.) 

### Position 6

```{r}
ggboard("XGID=-B-C-CC-A---a--bcd-eB-A---:0:0:1:43:2:4:0:7:10")
```

A technical play that White got wrong, by a surprising amount. White wanted to stay back, making it harder for Black to play safe, taking advantage of a stronger home board. The play made was 8/4 5/2, maintaing contact, trying to improve the home board fast.

The key is, that White is ahead in the race, not behind, and would like to release the back checker to steer into a more race-like position. 

Any good play involves 22/18 with the 4:

```{r}
"XG Roller++
1. 22/18 5/2      eq:-0,274
2. 22/18 8/5      eq:-0,274 (-0,001)
3. 22/18 6/3      eq:-0,311 (-0,037)
4. 6/2 5/2        eq:-0,429 (-0,155)
5. 8/4 5/2        eq:-0,459 (-0,186)
" %>% 
  cat()
```

After 22/18, Black's best rolls are the ones that hit on his 7-point without leaving returns: 61, 62, 21, 11, 22, 33 equalizing the race, along with 66 and 55, taking the lead in the race.

Black's *worst* roll, surprisingly, is the quiet 43: 13/10 6/2, the smallest roll that fails to hit, after which the game is more or less equal. Other below-average rolls are the ones that hit but leaves dangerous returns: 52, 51, 42, 41, 32, 64. Black can play safe with these rolls, but is better off taking a moderate risk to get ahead in the race.

After the actual 8/4 5/2, Black's worst roll is 51, played 8/2, voluntarily leaving only 4 indirect shots. The point is that the supposedly contact-maintaining play generates a lot fewer immediate shots than the contact-breaking play.

22/18 5/2 is so much better than 8/4 5/2 because it is better *both* for the race and for the hits.

In this variant, where White is behind in the race, 22/18 is still correct, although not by such a big margin, since it doesn't have the added racing benefits:

```{r}
#| layout-ncol: 2
ggboard("XGID=---C-CC-A--Ba--bcd-eB-A---:0:0:1:43:2:4:0:7:10")

"XG Roller++
1. 22/18 11/8        eq:-0,598
2. 22/18 8/5         eq:-0,605 (-0,007)
3. 22/18 5/2         eq:-0,608 (-0,011)
4. 11/4              eq:-0,623 (-0,025)
5. 22/18 6/3         eq:-0,626 (-0,028)" %>% 
  cat("\n", ., sep = "") 
```

## Conclusions

Examining some errors from this Milestone Tournament, the findings that surprised me the most where:

- I seem to *overestimate* some attacking positions
- I grossly *underestimate* how much more careful the leader should be to double in middle game positions with gammons, when leading something like 3-away versus 5-away or 7-away. Likely because I failed to pay attention how impactful potential recubes are at these scores.
- I tend to miss plays where making the ace-point is correct.
- Some technical positions can be deceptively tricky. 

I think I should dedicate another note to some of these themes.