---
title: "Milestones III, 2023"
author: "Lasse Hjorth Madsen"
date: today 
format: 
  html:
    fig-width: 4
    fig-height: 4
toc: true
toc-expand: true
editor: source
execute:
  echo: false
  warning: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(kableExtra)

options(scipen = 9999)
devtools::load_all(path = "../../backgammon")

met_file <- devtools::package_file("data-raw/Kazaross XG2.met", path = "../../backgammon")
met <- get_met(filename = met_file)
```

## Why this?

This is a collection of positions from the Milestones III tournament in Denmark, 2023, where players compete for wins and for best performance rating (PR). Every player gets to play 15 7-point matches. I will collect about one problem from each of my own matches: Positions where I made some mistake, that I found difficult, surprising or otherwise interesting.

In some cases, I show also variations, to better understand what is going on. I will organize the positions only into cube actions and checker plays, adding more as they come.

## Cube actions

### Position 1

```{r}
ggboard("XGID=bA-a-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:2:0:0:7:10")
```

Over the board, this looked to me like a strong double. White has an excellent forward structure, a fat racing lead, and a 4-point prime that can be extended from the back in relative safety, since Black has two checkers on the bar. The only weakness is White's blot on the ace point, but getting hit is not a big deal, so it can't be much of a liability.

Meanwhile, Black has made almost no progress; only the 11-point is made.

My guess was a strong double, and a smallish take for Black. It is, however, the other way around: White has only a small double, Black has a huge take:

```{r}
"Cubeful Equities:
       No double:     +0,690 (-0,047)
       Double/Take:   +0,736
       Double/Pass:   +1,000 (+0,264)
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller" %>%
cat()
```

Looking at the outcome probabilities, we see that White does have a big gammon threat, but also that Black wins more than a third of the games:

```{r}
# Analyzed in Rollout
# No double
#   Player Winning Chances:   65,12% (G:35,26% B:1,62%)
#   Opponent Winning Chances: 34,88% (G:7,79% B:0,49%)
# Double/Take
#   Player Winning Chances:   65,58% (G:36,41% B:2,44%)
#   Opponent Winning Chances: 34,42% (G:7,92% B:0,75%)

probs <- outcome_probs(c(65.58, 36.41, 2.44, 34.42, 7.92, 0.75)) %>% 
  `/`(100) %>% 
  round(3)

probs_fliped <- c(probs[4:6], probs[1:3])

probs %>% 
  probs_table() %>% 
  kable("html") %>%
  column_spec(2:5, width = "3.1cm")
```

One thing that I might have overlooked here, is that White has very few real crushing frequencies, because he's a little short on ammo. Black will almost always make some kind of anchor, and then have some kind of counter play agains White's three back checkers.

To see if this result is match-score related, let's check the same position for money, and while we're at it, also check if XG Roller ++ (fast, truncated rollouts), gets it right, compared to a full rollout at strong settings.

```{r}
#| layout-ncol: 2

ggboard("XGID=bA-a-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Cubeful Equities:
       No double:     +0,800 (-0,120)
       Double/Take:   +0,920
       Double/Pass:   +1,000 (+0,080)
1296 Games rolled with Variance Reduction.
Moves: 3-ply, cube decisions: XG Roller

Analyzed in XG Roller++
       No double:     +0,784 (-0,134)
       Double/Take:   +0,918
       Double/Pass:   +1,000 (+0,082)
" %>%
cat("\n", ., sep = "") 
```

For money, the double is bigger, and the take is smaller, so more in line with my first impression. It appears that the 2-point lead for White does make a difference.

But why?

The next table shows Black's takepoints under different assumption: No cube and no gammons (like a last-roll position); no cube and gammon rates from the table above; and finally taking both cube and gammons into account.

```{r}
tp_table <- tibble(
  `Cube assumptions` = c("Dead cube, no gammons", "Dead cube, gammons", "Both cube and gammons"),
  `Moneygame takepoint` = c(tp_money(c(0.5, 0, 0, 0.5, 0, 0), 0),
                            tp_money(probs_fliped, x = 0),
                            tp_money(probs_fliped, x = 0.68)
                            ),
  `7-away, 5-away takepoint` = c(tp(7, 5, 1, met, last_roll = TRUE),
                                 tp_info(7, 5, probs_fliped, 1, met)["tp_dead"],
                                 tp_info(7, 5, probs_fliped, 1, met)["tp_real"])
  )

tp_table %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `7-away, 5-away takepoint` - `Moneygame takepoint`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

The last-roll takepoint is not too different from money at this score, and it also doesn't make much difference to factor in gammons. When we consider possible recubes, however, we see that Black's takepoint at the score, is somewhat lower than it would be for money.

This is because he'll redouble more aggressively at the score. We can see why, by looking at *White's* takepoints for a potential recube, assuming the same gammon frequencies:

```{r}
tp_table <- tibble(
  `Cube assumptions` = c("Dead cube, no gammons", "Dead cube, gammons", "Both cube and gammons"),
  `Moneygame takepoint` = c(tp_money(c(0.5, 0, 0, 0.5, 0, 0), 0),
                            tp_money(probs, x = 0),
                            tp_money(probs, x = 0.68)
                            ),
  `5-away, 7-away recube takepoint` = c(tp(5, 7, 2, met, last_roll = TRUE),
                                 tp_info(5, 7, probs, 2, met)["tp_dead"],
                                 tp_info(5, 7, probs, 2, met)["tp_real"])
  )

tp_table %>% 
  mutate(across(where(is.numeric), round, 3),
         Difference = `5-away, 7-away recube takepoint` - `Moneygame takepoint`) %>% 
  kable("html") %>% 
  column_spec(2:4, width = "3.5cm")
```

So, it looks like a good part of my flawed evaluation of the original position, comes from the fact that Black can take somewhat more aggressively at the score; hence White needs more to have a really strong double.

The reason for Black's lower takepoint at the score, lies mostly in the potential redouble in positions where gammons are possible: It's hard for White to accept a 4-cube at 5-away 7-away, where a gammon loss means is much more costly than a gammon win.

Still, the position is also a clear take for money, so there's some reconsideration to do in that department too. What will it take to turn the position into a really tough take for money?

If we activate the awkward spare on the ace-point, we suddenly have a big pass. Perhaps not surprising, since the liability turns into a big asset. But still:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BCBB-A-cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
       No double:     +0,979 (-0,021)
       Double/Take:   +1,273 (+0,273)
       Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

Even if we put the spare checker in a less ideal place, we still get a double/pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BBBB-A-cCb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
       No double:     +0,929 (-0,071)
       Double/Take:   +1,045 (+0,045)
       Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

That spare has to move out of the action, to get to a marginal pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BBBB-A-cBb-Ac-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
       No double:     +0,913 (-0,087)
       Double/Take:   +1,011 (+0,011)
       Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

Finally, we get to a decent take, noting that Black is no longer far behind in the race:

```{r}
#| layout-ncol: 2
ggboard("XGID=b--a-BBBB-A-cBb--c-dB-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
       No double:     +0,874 (-0,072)
       Double/Take:   +0,947
       Double/Pass:   +1,000 (+0,053)" %>%
cat("\n", ., sep = "") 
```

One way to see how bad White's ace-point blot is, is to consider the original position after something like 31: 24/20. Now, most aces, like 21, 31, 41, and 51, are below-average rolls for Black, even though they hit. White is *hoping* to get hit (except with 41 and 11, trying for a high anchor); that's how bad the blot is.

Another experiment would be to improve White's opportunity to extend the prime from the back. A one-pip change can be huge: We go from a clear take to a pretty clear pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=bA-a-BBBBA--cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
       No double:     +0,856 (-0,144)
       Double/Take:   +1,059 (+0,059)
       Double/Pass:   +1,000
" %>%
cat("\n", ., sep = "") 
```

Another change in the same spirit is to bring in another builder for that important 9-point. It has about the samme effect as slotting:

```{r}
#| layout-ncol: 2
ggboard("XGID=bA-a-BBBB-A-cBbA-c-d--A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
       No double:     +0,854 (-0,146)
       Double/Take:   +1,065 (+0,065)
       Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

A final one-pip change that has a rather dramatic effect, is to just semi-active the ace-point checker so it easier becomes part of a prime, and is both easier and more attractive to cover. Again, we go from a clear take to a clear pass:

```{r}
#| layout-ncol: 2
ggboard("XGID=b-Aa-BBBB-A-cBb--c-dA-A-A-:0:0:1:00:0:0:0:0:10")

"Analyzed in XG Roller++
Cubeful Equities:
       No double:     +0,849 (-0,151)
       Double/Take:   +1,062 (+0,062)
       Double/Pass:   +1,000" %>%
cat("\n", ., sep = "") 
```

### Conclusion

-   A two-point lead in a 7-point match matters, especially if you might face a recube when gammons are still possible.
-   What seems to be a strong positional advantage, might not be so strong if there are also liabilities.
-   An awkward ace-point blot can be a rather big liability
-   Small things matter: One pip can make a it much easier to turn a strong position into a winning position.
